### 2024-01-26 demo.qmd
Add demo

---
title: Demonstration of matching route networks
format: gfm
---

```{r}
library(stplanr)
library(sf)
rnet_y = route_network_small["flow"]
rnet_y$id = 1:nrow(rnet_y)
# The target object
rnet_x = rnet_subset(osm_net_example[1], rnet_y)
# Get start and end points for each linestring in rnet_x:
rnet_x_s = lwgeom::st_startpoint(rnet_x)
rnet_x_e = lwgeom::st_endpoint(rnet_x)
plot(rnet_x$geometry, lwd = 9, col = "lightgrey")
plot(rnet_y["flow"], add = TRUE, lwd = 5)
plot(rnet_x_s, add = TRUE, pch = 3)
plot(rnet_x_e, add = TRUE, pch = 3)
```

The stplanr solution was as follows:

```{r}
rnet_y$quietness <- rnorm(nrow(rnet_y))
funs <- list(flow = sum, quietness = mean)
rnet_merged <- rnet_merge(rnet_x[1], rnet_y[c("flow", "quietness")],
  dist = 9, segment_length = 20, funs = funs
)
plot(rnet_y$geometry, lwd = 5, col = "lightgrey")
plot(rnet_merged["flow"], add = TRUE, lwd = 2)
```

Note that this leaves gaps in the network.
Also, the join syntax is a bit clunky.

Let's try doing the same thing but with `geos`.
First we'll reproject the data.

```{r}
rnet_x_projected = st_transform(rnet_x, "EPSG:27700")
rnet_y_projected = st_transform(rnet_y, "EPSG:27700")
```

Then we'll convert to `geos` objects.

```{r}
rnet_x_geos = geos::as_geos_geometry(rnet_x_projected)
rnet_y_geos = geos::as_geos_geometry(rnet_y_projected)
```

We'll create a buffer around `rnet_x`:

```{r}
params = geos::geos_buffer_params(end_cap_style = "flat")
rnet_x_buffer = geos::geos_buffer(rnet_x_geos, 9, params = params)
plot(rnet_x_buffer)
plot(rnet_y_geos, add = TRUE, col = "red", lwd = 2)
```

Now let's 'chop' the source geometry into segments that fit within the buffer:

```{r}
rnet_xbl = geos::geos_boundary(rnet_x_buffer)
rnet_xblb = geos::geos_buffer(rnet_xbl, 3, params = params)
rnet_xlbc = geos::geos_make_collection(rnet_xblb)
rnet_xlbcu = geos::geos_unary_union(rnet_xlbc)
rnet_y_chopped = geos::geos_intersection(
  rnet_y_geos,
  rnet_xlbcu
)
plot(rnet_xlbcu, col = "lightgrey")
```

```{r}
plot(rnet_xlbcu, col = "lightgrey")
plot(rnet_y_chopped, add = TRUE, col = "red", lwd = 2)
```

The red bits are the parts of the source geometry `rnet_y` that we *don't* want.
Let's get the bits that we *do* want:

```{r}
rnet_y_chopped = geos::geos_difference(
  rnet_y_geos,
  rnet_xlbcu
)
plot(rnet_xlbcu)
plot(rnet_x_buffer, add = TRUE, col = "lightgrey", border = NA)
plot(rnet_y_chopped, add = TRUE, col = "red", lwd = 2)
```

For every 'chopped' linestring there is at least one matching linestring in `rnet_y`.
Let's find them as follows:

```{r}
rnet_ycj = geos::geos_inner_join_keys(
  rnet_y_chopped,
  rnet_y_geos
)
plot(rnet_ycj)
```

That shows that we're not quite ready to join the data: we need to exclude the red bits that are outside the original buffer.

```{r}
#| echo: false
#| eval: false
multilinestring = geos::geos_unary_union(
    geos::geos_make_collection(c(
      geos::geos_make_linestring(1:3, 1:3),
      geos::geos_make_linestring(11:13, 11:13)
    ))
)
# This does not generate a linestring:
geos::geos_unnest(multilinestring)
# # Convert back to linestring with ???:
# multilinestring_sf = sf::st_as_sf(multilinestring)
# sf::st_cast(multilinestring_sf, "LINESTRING")
geos::geos_unnest(multilinestring, keep_multi = FALSE)
```

```{r}
rnet_ycl = geos::geos_unnest(rnet_y_chopped, keep_multi = FALSE)
rnet_ycj = geos::geos_inner_join_keys(
  rnet_ycl,
  rnet_y_geos
)
plot(rnet_ycj)
```

We can also join `rnet_y_chopped` and `rnet_ycl` to `rnet_x_buffer` to get the buffer geometry:

```{r}
rnet_ycj = geos::geos_inner_join_keys(
  rnet_y_chopped,
  rnet_x_buffer
)
nrow(rnet_ycj)
length(rnet_y_chopped)
length(unique(rnet_ycj$x))
length(unique(rnet_ycj$y))
plot(rnet_ycj)
rnet_y

rnet_yclj = geos::geos_inner_join_keys(
  rnet_ycl,
  rnet_x_buffer
)
nrow(rnet_yclj)
length(rnet_ycl)
plot(rnet_yclj)
rnet_y
```

Now let's join the dataframe versions:

```{r}
rnet_y_df = sf::st_drop_geometry(rnet_y)
rnet_y_df_expanded = rnet_y_df[rnet_ycj$x, ]
rnet_x_df = sf::st_drop_geometry(rnet_x)
rnet_x_df_expanded = rnet_x_df[rnet_ycj$y, ]
rnet_y_df_expanded$osm_id = rnet_x_df_expanded$osm_id
nrow(rnet_y_df_expanded)
nrow(rnet_x)
rnet_x_agg = rnet_y_df_expanded |>
  dplyr::group_by(osm_id) |>
  dplyr::summarise(
    flow = sum(flow),
    quietness = mean(quietness)
  )
rnet_x_joined = dplyr::left_join(
    rnet_x,
    rnet_x_agg,
)
plot(rnet_x_joined["flow"])
```

Let's compare the old and new joined flows:

```{r}
# place plots side by side
par(mfrow = c(1, 2))
plot(rnet_y$geometry, lwd = 5, col = "lightgrey")
plot(rnet_merged["flow"], add = TRUE, lwd = 2)

plot(rnet_y$geometry, lwd = 5, col = "lightgrey")
plot(rnet_x_joined["flow"], add = TRUE, lwd = 2)
par(mfrow = c(1, 1))
```

To package this up we used the following commands:

```{r}
#| eval: false
usethis::use_description()
usethis::use_package("geos")
```

We created a function that takes `rnet_x` and `rnet_y` and returns a dataframe with the joined data:

```{r}
rnet_x = rnet_x_projected
rnet_y = rnet_y_projected
distance = 9
dist_chop = 0.1
rnet_join = function(
    rnet_x,
    rnet_y,
    distance = 9,
    dist_chop = 0.1
    ) {
  # Check CRS is projected:
  stopifnot(
    sf::st_is_longlat(rnet_x) == FALSE,
    sf::st_is_longlat(rnet_y) == FALSE
  )
  rnet_x_geos = geos::as_geos_geometry(rnet_x)
  rnet_y_geos = geos::as_geos_geometry(rnet_y)
  params = geos::geos_buffer_params(end_cap_style = "flat")
  rnet_x_buffer = geos::geos_buffer(rnet_x_geos, distance, params = params)
  rnet_xbl = geos::geos_boundary(rnet_x_buffer)
  rnet_xblb = geos::geos_buffer(rnet_xbl, dist_chop, params = params)
  rnet_xlbc = geos::geos_make_collection(rnet_xblb)
  rnet_xlbcu = geos::geos_unary_union(rnet_xlbc)
  rnet_y_chopped = geos::geos_difference(
    rnet_y_geos,
    rnet_xlbcu
  )
  rnet_ycj = geos::geos_inner_join_keys(
    rnet_y_chopped,
    rnet_x_buffer
  )
  rnet_y_df = sf::st_drop_geometry(rnet_y)
  rnet_y_df_expanded = rnet_y_df[rnet_ycj$x, ]
  rnet_y_df_expanded$id_x = rnet_x_df[[1]][rnet_ycj$y]
  # Replace 'id_x' with the name of the first column in rnet_x_df:
  nx = which(names(rnet_y_df_expanded) == "id_x")
  new_name = names(rnet_x_df)[1]
  names(rnet_y_df_expanded)[nx] = new_name
  rnet_y_df_expanded
}
# Test it:
rnet_join(rnet_x, rnet_y)
```

### 2024-01-26 demo.qmd
Update with osm ids

@@
 # Test it:
-rnet_join(rnet_x, rnet_y)
+res = rnet_join(rnet_x, rnet_y)
+res
+res |>
+  dplyr::group_by(id) |>
+  dplyr::summarise(
+    osm_ids = paste(osm_id, collapse = ", ")
+  )
+
 ```

### 2024-01-26 demo.qmd
Add benchmark

@@
 ```{r}
+# Convert to linestring:
 rnet_xbl = geos::geos_boundary(rnet_x_buffer)
@@
 rnet_xlbcu = geos::geos_unary_union(rnet_xlbc)
-rnet_y_chopped = geos::geos_intersection(
+rnet_y_remove = geos::geos_intersection(
   rnet_y_geos,
@@
 plot(rnet_xlbcu, col = "lightgrey")
-plot(rnet_y_chopped, add = TRUE, col = "red", lwd = 2)
+plot(rnet_y_remove, add = TRUE, col = "red", lwd = 2)
 ```
@@
 
-For every 'chopped' linestring there is at least one matching linestring in `rnet_y`.
-Let's find them as follows:
+<!-- For every 'chopped' linestring there is at least one matching linestring in `rnet_y`.
+Let's find them as follows: -->
 
 ```{r}
+#| echo: false
+#| eval: false
 rnet_ycj = geos::geos_inner_join_keys(
@@
 
-That shows that we're not quite ready to join the data: we need to exclude the red bits that are outside the original buffer.
+<!-- That shows that we're not quite ready to join the data: we need to exclude the red bits that are outside the original buffer. -->
 
@@
 rnet_y
+```
+
+```{r}
+rnet_ycj2 = geos::geos_inner_join_keys(
+  rnet_y_geos,
+  rnet_x_buffer
+)
+waldo::compare(rnet_ycj, rnet_ycj2)
+```
+
+```{r}
 ```
@@
 ```
+
+```{r}
+#| eval: false
+#| echo: false
+# Calculate angle:
+geos::
+```
+
+# Benchmark
+
+Let's test it on a larger dataset:
+
+```{r}
+rnet_x = sf::read_sf("https://github.com/ropensci/stplanr/releases/download/v1.0.2/rnet_x_ed.geojson")
+rnet_y = sf::read_sf("https://github.com/ropensci/stplanr/releases/download/v1.0.2/rnet_y_ed.geojson")
+```
+
+```{r}
+plot(rnet_x$geometry, lwd = 9, col = "lightgrey")
+plot(rnet_y["value"], add = TRUE, lwd = 5)
+plot(rnet_x_s, add = TRUE, pch = 3)
+plot(rnet_x_e, add = TRUE, pch = 3)
+```
+
+Let's compare the `stplanr` implementation with the new implementation:
+
+```{r}
+# Stplanr implementation:
+system.time({
+rnet_merged = rnet_merge(rnet_x, rnet_y["value"], dist = 9, segment_length = 20, funs = list(value = sum))
+})
+plot(rnet_merged["value"], lwd = 3)
+```
+
+```{r}
+devtools::load_all()
+rnet_xp = rnet_x |>
+  sf::st_transform("EPSG:27700")
+rnet_yp = rnet_y |>
+  sf::st_transform("EPSG:27700") 
+system.time({
+rnet_matched = rnet_match(rnet_xp, rnet_yp, distance = 9, dist_chop = 1)
+rnet_matched_agg = rnet_matched |>
+  dplyr::group_by(identifier) |>
+  dplyr::summarise(value = sum(value))
+rnet_joined = rnet_x |>
+  dplyr::left_join(rnet_matched_agg)
+})
+summary(rnet_joined)
+# rnet_joined$value
+plot(rnet_joined["value"], lwd = 3)
+```
+
+```{r}
+```

### 2024-01-26 R/rnet_match.R

@@ -1,0 +1,62 @@
+rnet_match = function(
+    rnet_x,
+    rnet_y,
+    distance = 9,
+    dist_chop = 0.1,
+    rnet_y_df = NULL,
+    rnet_x_df = NULL) {
+  if (is(rnet_x, "sf") && is(rnet_y, "sf")) {
+    rnet_x_geos = geos::as_geos_geometry(rnet_x)
+    rnet_y_geos = geos::as_geos_geometry(rnet_y)
+    # stopifnot(
+    #   sf::st_is_longlat(rnet_x) == FALSE,
+    #   sf::st_is_longlat(rnet_y) == FALSE
+    # )
+  } else if (is(rnet_x, "geos_geometry") && is(rnet_y, "geos_geometry")) {
+    rnet_x_geos = rnet_x
+    rnet_y_geos = rnet_y
+  } else {
+    stop("rnet_x and rnet_y must be sf objects or geos_geometry objects")
+  }
+  rnet_ycj = rnet_match_geos(
+    rnet_x_geos,
+    rnet_y_geos,
+    distance = distance,
+    dist_chop = dist_chop
+  )
+  if (is.null(rnet_y_df)) {
+    rnet_y_df = sf::st_drop_geometry(rnet_y)
+  }
+  if (is.null(rnet_x_df)) {
+    rnet_x_df = sf::st_drop_geometry(rnet_x)
+  }
+  rnet_y_df_expanded = rnet_y_df[rnet_ycj$x, ]
+  rnet_y_df_expanded$id_x = rnet_x_df[[1]][rnet_ycj$y]
+  # Replace 'id_x' with the name of the first column in rnet_x_df:
+  nx = which(names(rnet_y_df_expanded) == "id_x")
+  new_name = names(rnet_x_df)[1]
+  names(rnet_y_df_expanded)[nx] = new_name
+  rnet_y_df_expanded
+}
+
+rnet_match_geos = function(
+    rnet_x_geos,
+    rnet_y_geos,
+    distance = 9,
+    dist_chop = 0.1) {
+  params = geos::geos_buffer_params(end_cap_style = "flat")
+  rnet_x_buffer = geos::geos_buffer(rnet_x_geos, distance, params = params)
+  rnet_xbl = geos::geos_boundary(rnet_x_buffer)
+  rnet_xblb = geos::geos_buffer(rnet_xbl, dist_chop, params = params)
[…]

### 2024-01-27 demo.qmd
Add rsgeo benchmark

@@
 ```{r}
+plot(rnet_y["value"], lwd = 5)
+```
+
+```{r}
 plot(rnet_x$geometry, lwd = 9, col = "lightgrey")
@@
 
+# With rsgeo
+
 ```{r}
+remotes::install_github("josiahparry/rsgeo", ref = "rent")
+
+rnet_x = sf::read_sf("https://github.com/ropensci/stplanr/releases/download/v1.0.2/rnet_x_ed.geojson") |> 
+  sf::st_transform(27700)
+rnet_y = sf::read_sf("https://github.com/ropensci/stplanr/releases/download/v1.0.2/rnet_y_ed.geojson") |> 
+  sf::st_transform(27700)
+
+plot(rnet_x)
+library(rsgeo)
+
+x <- as_rsgeo(rnet_x)
+y <- as_rsgeo(rnet_y)
+
+# 3rd arg is distance 4th is slope tolerance
+# distance is in units of the input data
+system.time({
+res <- rsgeo:::rnet_merge(x, y, 20, 0.2)
+
+excluded <- setdiff(1:length(x), res$from)
+
+from <- rnet_x[res$from,]
+to <- lapply(res$to, \(.i) rnet_y[.i,])
+to_mean_value <- sapply(res$to, \(.i) mean(rnet_y$value[.i]))
+
+rnet_matched_rsgeo = dplyr::bind_cols(from, value = to_mean_value)
+})
+
+plot(rnet_matched_rsgeo["value"], lwd = 3)
+
 ```
+
+```{r}
+#| eval: false
+#| echo: false
+# to be joined on
+plot(x[res$from], lwd = 2)
+# joined onto x from y
+plot(y[unique(unlist(res$to))], col = "red", lty = 3, add = TRUE, lwd = 2)
+# not used
+plot(x[excluded], col = "blue", lty = 3, add = TRUE)
+```
+
+Let's try matching the data:
+
+```{r}
+
+```

### 2024-01-27 demo.qmd
Add another rsgeo test

@@
 Also, the join syntax is a bit clunky.
+
+## With `{rsgeo}`
+
+Let's try with the `{rsgeo}` implementation:
+
+```{r}
+library(rsgeo)
+rnet_x = rnet_x |>
+  sf::st_transform("EPSG:27700")
+rnet_y = rnet_y |>
+  sf::st_transform("EPSG:27700")
+rnet_y$value = rnet_y$flow
+x <- as_rsgeo(rnet_x)
+y <- as_rsgeo(rnet_y)
+res <- rsgeo:::rnet_merge(x, y, 20, 0.2)  
+excluded <- setdiff(1:length(x), res$from)
+from <- rnet_x[res$from,]
+to <- lapply(res$to, \(.i) rnet_y[.i,])
+to_mean_value <- sapply(res$to, \(.i) mean(rnet_y$value[.i]))
+rnet_matched_rsgeo = dplyr::bind_cols(from, value = to_mean_value)
+plot(rnet_matched_rsgeo["value"], lwd = 3)
+```
+
+## With `{geos}`
 

### 2024-01-29 r/R/extendr-wrappers.R
restructure and add rust lib

@@ -1,0 +1,19 @@
+# Generated by extendr: Do not edit by hand
+
+# nolint start
+
+#
+# This file was created with the following call:
+#   .Call("wrap__make_rnetmatch_wrappers", use_symbols = TRUE, package_name = "rnetmatch")
+
+#' @docType package
+#' @usage NULL
+#' @useDynLib rnetmatch, .registration = TRUE
+NULL
+
+rnet_match_one_tree <- function(x, y, distance_tolerance, slope_tolerance, is_projected) .Call(wrap__rnet_match_one_tree, x, y, distance_tolerance, slope_tolerance, is_projected)
+
+rnet_match_two_trees <- function(x, y, distance_tolerance, slope_tolerance, is_projected) .Call(wrap__rnet_match_two_trees, x, y, distance_tolerance, slope_tolerance, is_projected)
+
+
+# nolint end

### 2024-01-29 r/R/rnet_match.R

@@ -1,0 +1,45 @@
+#' Match two road networks
+#'
+#' @details
+#'
+#' `x` and `y` are cast as geoarrow arrays via `geoarrow::as_geoarrow_array()`.
+#'
+#' Spherical geometries are not supported at this moment. Using them will result
+#' in inaccurate measurements.
+#'
+#' With a smaller number of features, building a spatial index on only the `x`
+#' geometries will be faster than building a spatial index on both `x` and `y`.
+#' However, at a larger number of features in `y`, building a spatial index
+#' can result in significant performance enhancements.
+#'
+#' @param x the target of the join
+#' @param y the features that will be joined to `x`
+#' @param dist_tolerance the maximum distances that each line segment from `y` can be away from `x`
+#' @param slope_tolerance the maximum difference in slope between line segments to be considered a match
+#' @param trees whether to build a spatial index on `x` or `x` and `y`.
+#' @export
+rnet_match <- function(x, y, dist_tolerance, slope_tolerance, trees = c("xy", "x")) {
+
+  trees <- match.arg(trees, several.ok = FALSE)
+
+  # if x or y are sf objects extract geometry
+  if (inherits(x, "sf")) {
+    x <- sf::st_geometry(x)
+  }
+
+  if (inherits(y, "sf")) {
+    y <- sf::st_geometry(y)
+  }
+
+  # TODO: handle other geometry types (geos & rsgeo)
+  f <- switch(trees, x = rnet_match_one_tree, y = rnet_match_two_trees)
+
+  f(
+    geoarrow::as_geoarrow_array(x),
+    geoarrow::as_geoarrow_array(y),
+    dist_tolerance,
+    slope_tolerance,
+    TRUE
+  )
+}
+

### 2024-01-29 rust/src/overlap.rs

@@ -1,0 +1,74 @@
+use geo_types::Rect;
+use std::ops::Range;
+
+// TODO for handling geographic CRS
+// calculate the distance from the top left to the bottom left corners
+pub fn x_range(rect: &Rect) -> Range<f64> {
+    rect.min().x..rect.max().x
+}
+
+// TODO for handling geographic CRS
+// calculate the distance from the top left to the top right corners
+pub fn y_range(rect: &Rect) -> Range<f64> {
+    rect.min().y..rect.max().y
+}
+
+pub fn overlap_range(r1: Range<f64>, r2: Range<f64>) -> Option<Range<f64>> {
+    if r1.end < r2.start || r2.end < r1.start {
+        None
+    } else {
+        Some(r1.start.max(r2.start)..r1.end.min(r2.end))
+    }
+}
+
+// Given the overlap in the domain and range
+// we can calculate the segment lenth of the line that is provided
+// we use the bounding box as a &Rect to determine the width or height
+// of the triangle
+// TODO: solve_segment_length does not handle geographic CRS yet 
+// TO support geographic CRS overlap_range() needs to be able to calculate haversine distance for the overlap
+// this will probably need to be handle by x_range and y_range?
+pub fn solve_segment_length(
+    x_overlap: Option<Range<f64>>,
+    y_overlap: Option<Range<f64>>,
+    bbox: &Rect,
+) -> f64 {
+    if x_overlap.is_some() && y_overlap.is_some() {
+        let (base_w, base_h) = wh(&bbox);
+        let dy = solve_dy(y_overlap.unwrap());
+        let dx = solve_dx(dy, base_w, base_h);
+        solve_h(dx, dy)
+    } else if x_overlap.is_some() {
+        let x_over = x_overlap.unwrap();
+        x_over.end - x_over.start
+    } else if y_overlap.is_some() {
+        let y_over = y_overlap.unwrap();
+        y_over.end - y_over.start
+    } else {
+        unreachable!() // this should never happen
+    }
+}
+
+// get height and width from a Line
+// do this by passing in the bounding rectangle
+// (width, height)
+fn wh(x: &Rect) -> (f64, f64) {
+    let (x1, y1) = x.min().x_y();
+    let (x2, y2) = x.max().x_y();
+    (x2 - x1, y2 - y1)
+}
+
+// Solve for dy:
[…]

### 2024-01-29 rust/src/structs.rs

@@ -1,0 +1,61 @@
+use geo::{BoundingRect, EuclideanDistance, HaversineDistance};
+use geo::{Line, Point};
+use rstar::RTreeObject;
+use rstar::AABB;
+
+
+pub enum CrsType {
+    Projected,
+    Geographic,
+}
+
+/// Custom struct to be used to insert into RTree
+/// Represents a component `Line` of a target `LineString`.
+/// The tuple stores the `Line` struct and the distance buffer to be used.
+/// It's [rstar::Envelope] method grows the [rstar::AABB] in x and y directions
+/// by the distance.
+pub struct TarLine(pub Line<f64>, pub f64);
+impl TarLine {
+    /// Create an AABB from the contained `Line`
+    pub fn envelope(&self) -> AABB<Point> {
+        let half_d = self.1 / 2.0;
+        let bb = self.0.bounding_rect();
+        let (ll_x, ll_y) = bb.min().x_y();
+        let (ur_x, ur_y) = bb.max().x_y();
+        let ll = Point::new(ll_x - half_d, ll_y - half_d);
+        let ur = Point::new(ur_x + half_d, ur_y + half_d);
+        AABB::from_corners(ll, ur)
+    }
+
+    /// Using geographic coordinate systems should be avoided with this algorithm.
+    /// Measuring distance in geographic space between two lines finds the minimum
+    /// distance between vertices whereas the euclidean distance between two lines
+    /// considers all possible distances.
+    /// Geographic distance may create false negatives.
+    pub fn dist_by_crs(&self, other: &Line, crs: &CrsType) -> f64 {
+        match crs {
+            CrsType::Projected => self.0.euclidean_distance(other),
+            CrsType::Geographic => {
+                let x = self.0;
+                x.start_point().haversine_distance(&other.start_point())
+                    .min(
+                        x.start_point().haversine_distance(&other.end_point())
+                    )
+                    .min(
+                        x.end_point().haversine_distance(&other.start_point())
+                    )
+                    .min(
+                        x.end_point().haversine_distance(&other.end_point())
+                    )
+            },
+
+        }
+    }
+}
+
+impl RTreeObject for TarLine {
+    type Envelope = AABB<Point>;
+    fn envelope(&self) -> Self::Envelope {
+        self.envelope()
+    }
+}

### 2024-01-29 r/README.qmd
Add placeholder readme for r bit

---
format: gfm
---

```{r}
devtools::install()
```

### 2024-01-29 r/README.qmd

@@
 ```
+
+For local development run this in the `r` directory:
+
+```{r}
+devtools::load_all()
+```
+
+```{r}
+library(dplyr)
+```
+
+## Example
+
+```{r}
+edinburgh_zones = zonebuilder::zb_zone("Edinburgh")
+edinburgh_central_3km = edinburgh_zones |>
+  filter(circle_id <= 2) |>
+  sf::st_union()
+osm_edinburgh_central = osmextract::oe_get_network("Edinburgh", boundary = edinburgh_central_3km, boundary_type = "clipsrc")
+meuse_lane = osm_edinburgh_central |>
+  filter(name == "Meuse Lane")
+meuse_lane_buffer = sf::st_buffer(meuse_lane, 100)
+osm_edinburgh_central = osm_edinburgh_central |>
+  sf::st_intersection(meuse_lane_buffer)
+osm_edinburgh_central |>
+  select(highway) |>
+  plot()
+```
+
+```{r}
+princes_street_central = osm_edinburgh_central |>
+  mutate(n = 1:n()) |>
+  filter(name == "Princes Street")
+princes_street_central |>
+  select(highway) |>
+  plot()
+# Check each link:
+princes_street_central |>
+  select(n) |>
+  plot(lwd = 5)
+```
+
+```{r}
+princes_street_minimal = princes_street_central |>
+  slice(1:4)
+princes_street_minimal = princes_street_minimal |>
+  mutate(value = 1:n()) |>
+  select(value, name, osm_id)
+princes_street_minimal |>
+  select(value) |>
+  plot(lwd = 5)
+```
+
+```{r}
+m = mapview::mapview(princes_street_minimal)
+princes_street_minimal_x = mapedit::editMap(m)
+names(princes_street_minimal_x)
+princes_street_minimal_x_1 = princes_street_minimal_x$drawn
+princes_street_minimal_x_1 = princes_street_minimal_x_1 |>
+  transmute(id = 1:n())
+```
+
+```{r}
+bind_rows(
+  princes_street_minimal |>
+    transmute(layer = "y"),
+  princes_street_minimal_x_1 |>
+    transmute(layer = "x")
+) |>
+  plot(lwd = 5)
+```
+
+```{r}
+dir.create("data-raw")
+dir.create("data-raw/geojson")
+sf::write_sf(princes_street_minimal, "data-raw/geojson/princes_street_minimal.geojson")
+sf::write_sf(princes_street_minimal_x_1, "data-raw/geojson/princes_street_minimal_x_1.geojson")
+```
+
+```{r}
+```

### 2024-01-29 r/README.qmd

@@
 ```{r}
+#| eval: false
 m = mapview::mapview(princes_street_minimal)
@@
 ```{r}
+#| echo: false
+#| eval: false
+princes_street_minimal_x_1_points = sf::st_cast(princes_street_minimal_x_1, "POINT")
+princes_street_minimal_x_1_point_buffers = sf::st_buffer(princes_street_minimal_x_1_points, 5) |>
+  sf::st_union()
+plot(princes_street_minimal_x_1_point_buffers)
+plot(princes_street_minimal_x_1, add = TRUE)
+```
+
+```{r}
+princes_street_minimal_x_1 = sf::st_difference(princes_street_minimal_x_1, princes_street_minimal_x_1_point_buffers)
+```
+
+```{r}
 bind_rows(
   princes_street_minimal |>
-    transmute(layer = "y"),
+    transmute(value = as.character(value)),
   princes_street_minimal_x_1 |>
-    transmute(layer = "x")
+    transmute(value = "x")
 ) |>
@@
 dir.create("data-raw/geojson")
-sf::write_sf(princes_street_minimal, "data-raw/geojson/princes_street_minimal.geojson")
-sf::write_sf(princes_street_minimal_x_1, "data-raw/geojson/princes_street_minimal_x_1.geojson")
+sf::write_sf(princes_street_minimal, "data-raw/geojson/princes_street_minimal.geojson", delete_dsn = TRUE)
+sf::write_sf(princes_street_minimal_x_1, "data-raw/geojson/princes_street_minimal_x_1.geojson", delete_dsn = TRUE)
 ```

### 2024-01-29 r/README.qmd
Update readme

--- 2024-01-29 r/README.qmd
@@
 ```
+
+For local development run this in the `r` directory:
+
+```{r}
+devtools::load_all()
+```
+
+```{r}
+library(dplyr)
+```
+
+Basic test
+
+```{r}
+x = sf::read_sf("data-raw/geojson/princes_street_minimal.geojson")
+y = sf::read_sf("data-raw/geojson/princes_street_minimal_x_1.geojson")
+```
+
+```{r}
+bind_rows(
+  princes_street_minimal |>
+    transmute(value = as.character(value)),
+  princes_street_minimal_x_1 |>
+    transmute(value = "x")
+) |>
+  plot(lwd = 5)
+```

### 2024-01-29 r/data-raw/generate-test-geojsons.qmd
Update readme

---
format: gfm
---

```{r}
devtools::install()
```

For local development run this in the `r` directory:

```{r}
devtools::load_all()
```

```{r}
library(dplyr)
```

## Example

```{r}
edinburgh_zones = zonebuilder::zb_zone("Edinburgh")
edinburgh_central_3km = edinburgh_zones |>
  filter(circle_id <= 2) |>
  sf::st_union()
osm_edinburgh_central = osmextract::oe_get_network("Edinburgh", boundary = edinburgh_central_3km, boundary_type = "clipsrc")
meuse_lane = osm_edinburgh_central |>
  filter(name == "Meuse Lane")
meuse_lane_buffer = sf::st_buffer(meuse_lane, 100)
osm_edinburgh_central = osm_edinburgh_central |>
  sf::st_intersection(meuse_lane_buffer)
osm_edinburgh_central |>
  select(highway) |>
  plot()
```

```{r}
princes_street_central = osm_edinburgh_central |>
  mutate(n = 1:n()) |>
  filter(name == "Princes Street")
princes_street_central |>
  select(highway) |>
  plot()
# Check each link:
princes_street_central |>
  select(n) |>
  plot(lwd = 5)
```

```{r}
princes_street_minimal = princes_street_central |>
  slice(1:4)
princes_street_minimal = princes_street_minimal |>
  mutate(value = 1:n()) |>
  select(value, name, osm_id)
princes_street_minimal |>
  select(value) |>
  plot(lwd = 5)
```

```{r}
#| eval: false
m = mapview::mapview(princes_street_minimal)
princes_street_minimal_x = mapedit::editMap(m)
names(princes_street_minimal_x)
princes_street_minimal_x_1 = princes_street_minimal_x$drawn
princes_street_minimal_x_1 = princes_street_minimal_x_1 |>
  transmute(id = 1:n())
```

```{r}
#| echo: false
#| eval: false
princes_street_minimal_x_1_points = sf::st_cast(princes_street_minimal_x_1, "POINT")
princes_street_minimal_x_1_point_buffers = sf::st_buffer(princes_street_minimal_x_1_points, 5) |>
  sf::st_union()
plot(princes_street_minimal_x_1_point_buffers)
plot(princes_street_minimal_x_1, add = TRUE)
```

```{r}
```

```{r}
```

```{r}
```

```{r}
bind_rows(
  princes_street_minimal |>
    transmute(value = as.character(value)),
  princes_street_minimal_x_1 |>
    transmute(value = "x")
) |>
  plot(lwd = 5)
```

```{r}
dir.create("geojson")
sf::write_sf(princes_street_minimal, "geojson/princes_street_minimal.geojson", delete_dsn = TRUE)
sf::write_sf(princes_street_minimal_x_1, "geojson/princes_street_minimal_x_1.geojson", delete_dsn = TRUE)
```

```{r}
```

### 2024-01-29 r/README.qmd
Add minimal example with plausible result

@@
 ```{r}
-x = sf::read_sf("data-raw/geojson/princes_street_minimal.geojson")
-y = sf::read_sf("data-raw/geojson/princes_street_minimal_x_1.geojson")
+x = sf::read_sf("data-raw/geojson/princes_street_minimal_x_1.geojson")
+x = x |> sf::st_transform("EPSG:27700")
+y = sf::read_sf("data-raw/geojson/princes_street_minimal.geojson")
+y = y |> sf::st_transform("EPSG:27700")
 ```
@@
 ```
+
+```{r}
+x_joined_df = rnet_match(x, y, dist_tolerance = 20, slope_tolerance = 0.1, trees = "xy")
+x_aggregated_x = x_joined_df |>
+  group_by(j) |>
+  mutate(ij = y$value * shared_len) |>
+  group_by(j) |>
+  summarise(value = sum(ij))
+
+```
+
+```{r}
+```

### 2024-01-30 rust/src/overlap.rs
update overlap calculation:

@@ -1,4 +1,4 @@
-use geo_types::Rect;
+use geo_types::{Line, Point, Rect};
 use std::ops::Range;
 
 // TODO for handling geographic CRS
@@ -49,10 +49,57 @@ pub fn solve_segment_length(
     }
 }
 
+
+// When x range is known but y range is not, we need to solve for start and end points 
+// of the line segment
+pub fn solve_no_y_overlap(x_overlap: Range<f64>, x: &Line, slope: &f64) -> (Point, Point) {
+    let (known_x, known_y) = x.points().0.x_y();
+    let b = known_y - (slope * known_x); // Corrected calculation of b
+
+    let y1 = (slope * x_overlap.start) + b;
+    let y2 = (slope * x_overlap.end) + b;
+    let p1 = Point::new(x_overlap.start, y1);
+    let p2 = Point::new(x_overlap.end, y2);
+    (p1, p2)  
+}
+
+pub fn solve_no_x_overlap(y_overlap: Range<f64>, x: &Line, slope: &f64) -> (Point, Point) {
+    let (known_x, known_y) = x.points().0.x_y();
+    let b = known_y - (slope * known_x); // Corrected calculation of b
+
+    let x1 = (y_overlap.start - b) / slope;
+    let x2 = (y_overlap.end - b) / slope;
+    let p1 = Point::new(x1, y_overlap.start);
+    let p2 = Point::new(x2, y_overlap.end);
+    (p1, p2)  
+}
+
+pub fn solve_known_overlaps(
+    x_overlap: Range<f64>, 
+    y_overlap: Range<f64>, 
+    xbb : &Rect) -> (Point, Point) {
+
+        let dy = solve_dy(y_overlap.clone());
+        // this is the width of the overlapping bbox
+        // let base_w = x_overlap.end - x_overlap.start;
+        // this is the heeight of the bbox around xi itself
+        // _not_ the bbox of the overlapping area
+        let (base_w, base_h) = wh(&xbb);
+        // this is the length of the line from the side of the bbox
+        // to the end of the line segment
+        let dx = solve_dx(dy, base_w, base_h);
+
+        let x1 = x_overlap.end - dx;
+        let p1 = Point::new(x1, y_overlap.start);
+        let p2 = Point::new(x_overlap.end, y_overlap.end);
+        (p1, p2)
+}
+
+
 // get height and width from a Line
[…]

### 2024-01-30 rust/src/lib.rs
update overlap calculation:

@@ -1,4 +1,5 @@
 use geo::BoundingRect;
+use geo::EuclideanDistance;
 use std::collections::BTreeMap;
 
 mod overlap;
@@ -85,12 +86,16 @@ pub fn find_candidates_one_tree(
             let ybb = li.bounding_rect();
             let candidates = source_tree.locate_in_envelope_intersecting(&envelope);
 
+
             candidates.for_each(|cx| {
                 let xbb = cx.geom().bounding_rect();
                 let (i, x_slope) = cx.data;
                 let y_slope = li.slope();
+                
                 // compare slopes:
                 let is_tolerant = (x_slope - y_slope).abs() < slope_tolerance;
+
+                
 
                 // if the slopes are within tolerance then we check for overlap
                 if is_tolerant {
@@ -107,7 +112,25 @@ pub fn find_candidates_one_tree(
                         // if distance is less than or equal to tolerance, add the key
                         if d <= distance_tolerance {
                             // solve_segment_length() will always calculate in R2 space
-                            let shared_len = solve_segment_length(x_overlap, y_overlap, &xbb);
+                            
+                            // let sl1 = solve_segment_length(x_overlap, y_overlap.clone(), &xbb);
+
+                            let (p1, p2) =  if x_overlap.is_some() && y_overlap.is_some() {
+                                solve_known_overlaps(
+                                    x_overlap.unwrap(), 
+                                    y_overlap.unwrap(), 
+                                    &xbb
+                                )
+                            } else if x_overlap.is_some() {
+                                solve_no_y_overlap(x_overlap.unwrap(), &cx.geom(), &x_slope)
+                            } else if y_overlap.is_some() {
+                                solve_no_x_overlap(y_overlap.unwrap(), &cx.geom(), &x_slope)
+                            } else {
+                                unreachable!()
+                            };
+                        
+                            let shared_len = p1.euclidean_distance(&p2);
+                            
                             // add 1 for R indexing
                             // ensures that no duplicates are inserted. Creates a new empty vector is needed
                             let entry = matches.entry((i + 1) as i32).or_insert_with(Vec::new);

### 2024-01-30 r/README.qmd
Update results

@@
 ---
-
-```{r}
-devtools::install()
-```
 
@@
 devtools::load_all()
+```
+
+You may also want to install it locally:
+
+```{r}
+#| eval: false
+devtools::install()
+```
+
+And clean up previous builds:
+
+```{r}
+#| eval: false
+rextendr::clean()
 ```
@@
 ```{r}
+#| label: fig-x-and-y
 bind_rows(
-  princes_street_minimal |>
-    transmute(value = as.character(value)),
-  princes_street_minimal_x_1 |>
-    transmute(value = "x")
+  x |>
+    transmute(value = "x"),
+  y |>
+    transmute(value = as.character(value))
 ) |>
@@
 ```{r}
-x_joined_df = rnet_match(x, y, dist_tolerance = 20, slope_tolerance = 0.1, trees = "xy")
-x_aggregated_x = x_joined_df |>
-  group_by(j) |>
-  mutate(ij = y$value * shared_len) |>
-  group_by(j) |>
-  summarise(value = sum(ij))
+matched_df = rnet_match(x, y, dist_tolerance = 20, slope_tolerance = 0.1, trees = "xy")
+y_to_match = y |>
+  transmute(j = 1:n(), value) |>
+  sf::st_drop_geometry() 
+matched_df_y = left_join(matched_df, y_to_match)
+x$length_x = sf::st_length(x) |>
+  # Drop units:
+  units::drop_units()
+x_to_match = x |>
+  transmute(i = 1:n(), length_x) |>
+  sf::st_drop_geometry()
+matched_df_xy = left_join(matched_df_y, x_to_match)
+x_formula = matched_df_xy |>
+  mutate(value_formula = value * shared_len / length_x)
+x_formula
+x_with_values = x_formula |>
+  group_by(id = i) |>
+  summarise(
+    value = sum(value_formula),
+    length_x = first(length_x)
+    )
+x_joined = left_join(x, x_with_values)
+```
 
+The total length travelled on each network can be calculated as follows:
+
+```{r}
+total_distance_x = sum(x_joined$value * x_joined$length_x)
+round(total_distance_x)
+```
+
+And for `y`:
+
+```{r}
+total_distance_y = sum(as.numeric(y$value * sf::st_length(y)))
+```
+
+We can post process the joined x network to get the total length travelled on each network:
+
+```{r}
+x_joined = x_joined |>
+  mutate(value = value * total_distance_y / total_distance_x)
+```
+
+That results in these values:
+
+```{r}
+round(x_joined$value)
+```
+
+```{r}
+#| label: fig-x_joined
+y |>
+  select(value) |>
+  plot(lwd = 5)
+x_joined |>
+  select(value) |>
+  plot(lwd = 5)
 ```
@@
 ```
+
+```{r}
+```

### 2024-01-30 rust/src/lib.rs
use degrees. improve segment length calculation

@@ -1,5 +1,5 @@
 use geo::BoundingRect;
-use geo::EuclideanDistance;
+use geo::{EuclideanDistance, HaversineDistance};
 use std::collections::BTreeMap;
 
 mod overlap;
@@ -14,13 +14,13 @@ pub use crate::trees::*;
 pub fn find_candidates(
     x: impl Iterator<Item = geo_types::LineString>,
     y: impl Iterator<Item = geo_types::LineString>,
-    dist: f64,
+    distance_tolerance: f64,
     slope_tolerance: f64,
     crs_type: CrsType,
 ) -> BTreeMap<i32, Vec<(i32, f64)>> {
     let mut matches: BTreeMap<i32, Vec<(i32, f64)>> = BTreeMap::new();
     let source_tree = create_source_rtree(x);
-    let target_tree = create_target_rtree(y, dist);
+    let target_tree = create_target_rtree(y, distance_tolerance);
     let candidates = source_tree.intersection_candidates_with_other_tree(&target_tree);
 
     candidates.for_each(|(cx, cy)| {
@@ -31,40 +31,69 @@ pub fn find_candidates(
         let (i, x_slope) = cx.data;
         let (j, y_slope) = cy.data;
 
-        // compare slopes:
-        // TODO handle vertical lines
-        // TODO use degrees or radians instead of calculated slope
-        let is_tolerant = (x_slope - y_slope).abs() < slope_tolerance;
 
-        // if the slopes are within tolerance then we check for overlap
-        if is_tolerant {
-            let x_overlap = overlap_range(x_range(&xbb), x_range(&ybb));
-            let y_overlap = overlap_range(y_range(&xbb), y_range(&ybb));
+                // convert calculated slopes to degrees
+                let x_deg = x_slope.atan().to_degrees();
+                let y_deg = y_slope.atan().to_degrees();
 
-            // if theres overlap then we do a distance based check
-            // following, check that they're within distance tolerance,
-            // if so, calculate the shared length
-            if x_overlap.is_some() || y_overlap.is_some() {
-                // calculate the distance from the line segment
-                // if its within our threshold we include it;
-                let d = cy.geom().dist_by_crs(&cx.geom(), &crs_type);
-                // if distance is less than or equal to tolerance, add the key
-                if d <= dist {
-                    let shared_len = solve_segment_length(x_overlap, y_overlap, &xbb);
-                    // TODO remove addition of 1 for R indexing
-                    // add 1 for R indexing
[…]

### 2024-01-30 rust/src/structs.rs

@@ -18,12 +18,12 @@ pub struct TarLine(pub Line<f64>, pub f64);
 impl TarLine {
     /// Create an AABB from the contained `Line`
     pub fn envelope(&self) -> AABB<Point> {
-        let half_d = self.1 / 2.0;
+        let d = self.1;
         let bb = self.0.bounding_rect();
         let (ll_x, ll_y) = bb.min().x_y();
         let (ur_x, ur_y) = bb.max().x_y();
-        let ll = Point::new(ll_x - half_d, ll_y - half_d);
-        let ur = Point::new(ur_x + half_d, ur_y + half_d);
+        let ll = Point::new(ll_x - d, ll_y - d);
+        let ur = Point::new(ur_x + d, ur_y + d);
         AABB::from_corners(ll, ur)
     }
 

### 2024-01-30 rust/src/lib.rs
use angle tolerance instead

@@ -15,7 +15,7 @@ pub fn find_candidates(
     x: impl Iterator<Item = geo_types::LineString>,
     y: impl Iterator<Item = geo_types::LineString>,
     distance_tolerance: f64,
-    slope_tolerance: f64,
+    angle_tolerance: f64,
     crs_type: CrsType,
 ) -> BTreeMap<i32, Vec<(i32, f64)>> {
     let mut matches: BTreeMap<i32, Vec<(i32, f64)>> = BTreeMap::new();
@@ -37,7 +37,7 @@ pub fn find_candidates(
                 let y_deg = y_slope.atan().to_degrees();
 
                 // compare slopes:
-                let is_tolerant = (x_deg - y_deg).abs() < slope_tolerance;
+                let is_tolerant = (x_deg - y_deg).abs() < angle_tolerance;
 
                 // if the slopes are within tolerance then we check for overlap
                 if is_tolerant {
@@ -102,7 +102,7 @@ pub fn find_candidates_one_tree(
     x: impl Iterator<Item = geo_types::LineString>,
     y: impl Iterator<Item = geo_types::LineString>,
     distance_tolerance: f64,
-    slope_tolerance: f64,
+    angle_tolerance: f64,
     crs_type: CrsType,
 ) -> BTreeMap<i32, Vec<(i32, f64)>> {
     let mut matches: BTreeMap<i32, Vec<(i32, f64)>> = BTreeMap::new();
@@ -127,7 +127,7 @@ pub fn find_candidates_one_tree(
 
 
                 // compare slopes:
-                let is_tolerant = (x_deg - y_deg).abs() < slope_tolerance;
+                let is_tolerant = (x_deg - y_deg).abs() < angle_tolerance;
 
                 
 

### 2024-01-30 rust/src/structs.rs
use longer var name

@@ -18,12 +18,12 @@ pub struct TarLine(pub Line<f64>, pub f64);
 impl TarLine {
     /// Create an AABB from the contained `Line`
     pub fn envelope(&self) -> AABB<Point> {
-        let d = self.1;
+        let padding_dist = self.1;
         let bb = self.0.bounding_rect();
         let (ll_x, ll_y) = bb.min().x_y();
         let (ur_x, ur_y) = bb.max().x_y();
-        let ll = Point::new(ll_x - d, ll_y - d);
-        let ur = Point::new(ur_x + d, ur_y + d);
+        let ll = Point::new(ll_x - padding_dist, ll_y - padding_dist);
+        let ur = Point::new(ur_x + padding_dist, ur_y + padding_dist);
         AABB::from_corners(ll, ur)
     }
 

### 2024-01-30 r/R/rnet_match.R
update doc name

@@ -15,10 +15,10 @@
 #' @param x the target of the join
 #' @param y the features that will be joined to `x`
 #' @param dist_tolerance the maximum distances that each line segment from `y` can be away from `x`
-#' @param slope_tolerance the maximum difference in slope between line segments to be considered a match
+#' @param angle_tolerance the maximum difference in slope between line segments to be considered a match
 #' @param trees whether to build a spatial index on `x` or `x` and `y`.
 #' @export
-rnet_match <- function(x, y, dist_tolerance, slope_tolerance, trees = c("xy", "x")) {
+rnet_match <- function(x, y, dist_tolerance, angle_tolerance, trees = c("xy", "x")) {
 
   trees <- match.arg(trees, several.ok = FALSE)
 
@@ -38,7 +38,7 @@ rnet_match <- function(x, y, dist_tolerance, slope_tolerance, trees = c("xy", "x
     geoarrow::as_geoarrow_array(x),
     geoarrow::as_geoarrow_array(y),
     dist_tolerance,
-    slope_tolerance,
+    angle_tolerance,
     TRUE
   )
 }

### 2024-01-30 r/tests/testthat.R
Add test (#10)

@@ -1,0 +1,12 @@
+# This file is part of the standard setup for testthat.
+# It is recommended that you do not modify it.
+#
+# Where should you do additional test configuration?
+# Learn more about the roles of various files in:
+# * https://r-pkgs.org/testing-design.html#sec-tests-files-overview
+# * https://testthat.r-lib.org/articles/special-files.html
+
+library(testthat)
+library(rnetmatch)
+
+test_check("rnetmatch")

### 2024-01-30 r/tests/testthat/test_matches.R

@@ -1,0 +1,9 @@
+# Aim: test that the matching works as expected
+
+test_that("matching works", {
+#   x = sf::read_sf(system.file("extdata", "y_negative.geojson", package = "rnetmatch"))
+  x = sf::read_sf("https://github.com/nptscot/rnetmatch/raw/main/r/data-raw/geojson/x_negative.geojson")
+  y = sf::read_sf("https://github.com/nptscot/rnetmatch/raw/main/r/data-raw/geojson/y_negative.geojson")
+  res = rnet_match(x, y, dist_tolerance = 1, slope_tolerance = 30)
+  expect_equal(round(res[1, 3]), 2)
+})

### 2024-01-30 r/tests/testthat/test_matches.R
Update r/tests/testthat/test_matches.R
Co-authored-by: Josiah Parry <josiah.parry@gmail.com>

@@ -4,6 +4,6 @@ test_that("matching works", {
 #   x = sf::read_sf(system.file("extdata", "y_negative.geojson", package = "rnetmatch"))
   x = sf::read_sf("https://github.com/nptscot/rnetmatch/raw/main/r/data-raw/geojson/x_negative.geojson")
   y = sf::read_sf("https://github.com/nptscot/rnetmatch/raw/main/r/data-raw/geojson/y_negative.geojson")
-  res = rnet_match(x, y, dist_tolerance = 1, slope_tolerance = 30)
+  res = rnet_match(x, y, dist_tolerance = 1, angle_tolerance = 30)
   expect_equal(round(res[1, 3]), 2)
 })

### 2024-01-30 r/data-raw/test-large-edinburgh.qmd
Add large example, close #13

---
format: gfm
---

```{r}
open_roads = sf::read_sf("https://github.com/nptscot/networkmerge/releases/download/v0.1/os_3km.geojson")
summary(duplicated(open_roads$geometry))
plot(open_roads$geometry)
open_roads = sf::st_make_valid(open_roads)
```

We can get some OSM data of the same area as follows:

```{r}
case_study_area = open_roads |>
  sf::st_union() |>
  sf::st_convex_hull() |>
  sf::st_transform("EPSG:4326")
osm_roads = osmextract::oe_get_network(
  "Edinburgh",
  mode = "driving",
  boundary = case_study_area,
  boundary_type = "clip"
)
plot(osm_roads$geometry)
```

```{r}
table(osm_roads$highway)
```

```{r}
osm_roads$value = 1
osm_roads$value[osm_roads$highway == "primary"] = 10
osm_roads$value[osm_roads$highway == "secondary"] = 9
osm_roads$value[osm_roads$highway == "tertiary"] = 8
osm_roads$value[osm_roads$highway == "unclassified"] = 7
osm_roads$value[osm_roads$highway == "residential"] = 2
```

```{r}
funs = list(value = sum)
```

```{r}
os_joined = stplanr::rnet_merge(
  open_roads |> dplyr::transmute(n = seq(nrow(open_roads))) |> sf::st_transform("EPSG:27700"),
  osm_roads |> sf::st_cast("LINESTRING") |> sf::st_zm() |> sf::st_transform("EPSG:27700"),
  funs = funs,
  dist = 20,
  segment_length = 10
  )
os_joined |>
  dplyr::select(value) |>
  plot(logz = TRUE)
```

### 2024-01-30 r/data-raw/test-large-edinburgh.qmd
Minor update

@@
   )
+plot(osm_roads$geometry, col = "grey", lwd = 3)
 os_joined |>
+  sf::st_transform("EPSG:4326") |>
   dplyr::select(value) |>
-  plot(logz = TRUE)
+  plot(logz = TRUE, add = TRUE)
+
 ```

### 2024-01-30 r/data-raw/test-large-edinburgh.qmd

@@
 ```
+```{r}
+x <- open_roads |> 
+  dplyr::transmute(n = seq(nrow(open_roads))) |> 
+  sf::st_transform("EPSG:27700") |> 
+  sf::st_cast("LINESTRING")
+
+y <-  osm_roads |> 
+  sf::st_cast("LINESTRING") |> 
+  sf::st_zm() |> 
+  sf::st_transform("EPSG:27700")
+
+res <- rnet_match(
+  x, y,
+  dist = 200,
+  angle_tolerance = 10,
+  trees = "xy"
+)
+
+head(res)
+```

### 2024-02-01 demo.qmd
using smaller segment_length

@@
 Note that this leaves gaps in the network.
+
+Although this gap issue can be fixed by using a smaller segment_length, e.g. segment_length = 5, this is not a computationally efficient way to solve the problem.
+
+```{r}
+rnet_y$quietness <- rnorm(nrow(rnet_y))
+funs <- list(flow = sum, quietness = mean)
+rnet_merged <- rnet_merge(rnet_x[1], rnet_y[c("flow", "quietness")],
+  dist = 9, segment_length = 5, funs = funs
+)
+plot(rnet_y$geometry, lwd = 5, col = "lightgrey")
+plot(rnet_merged["flow"], add = TRUE, lwd = 2)
+```
+
 Also, the join syntax is a bit clunky.

### 2024-02-01 demo.qmd
add intersection road example and show results of rnet_merge

@@
 plot(rnet_x_e, add = TRUE, pch = 3)
+```
+Here are the examples of an intersection road.
+
+```{r}
+rnet_intersection_complex = sf::read_sf("https://github.com/nptscot/rnetmatch/releases/download/intersection_road/intersection_example_complex.geojson")
+rnet_intersection_simple = sf::read_sf("https://github.com/nptscot/rnetmatch/releases/download/intersection_road/intersection_example_simple.geojson")
+rnet_intersection_simple$index = seq_along(rnet_intersection_simple$geometry)
+
+rnet_intersection_simple = rnet_intersection_simple[, c("index", setdiff(names(rnet_intersection_simple), "index"))]
+
+rnet_intersection_simple <- rnet_intersection_simple[c("index", "geometry")]
+
+plot(rnet_intersection_complex$geometry, lwd = 20, col = "blue")
+plot(rnet_intersection_simple$geometry, col = "red", add = TRUE, lwd = 5)
+```
+This is where rnet_merge funtion ard to handle the intersection road.
+```{r}
+rnet_xp = sf::st_transform(rnet_intersection_simple , "EPSG:27700")
+rnet_yp = sf::st_transform(rnet_intersection_complex, "EPSG:27700")
+names(rnet_xp)
+# Extract column names from the rnet_yp
+name_list = names(rnet_yp)
+
+# Initialize an empty list
+funs = list()
+
+# Loop through each name and assign it a function based on specific conditions
+for (name in name_list) {
+  if (name == "geometry") {
+    next  # Skip the current iteration
+  } else if (name %in% c("gradient", "quietness")) {
+    funs[[name]] = mean
+  } else {
+    funs[[name]] = sum
+  }
+}
+
+dist = 10
+angle = 5
+rnet_merged = stplanr::rnet_merge(rnet_xp, rnet_yp, dist = dist, funs = funs, max_angle_diff = angle, segment_lentgh = 10)
+
+library(tmap)
+tmap_mode("view")
+
+brks = c(0, 50, 100, 200, 500,1000,2000)
+m1 = tm_shape(rnet_intersection_complex) + tm_lines("commute_fastest_bicycle_go_dutch", palette = "viridis", breaks = brks)
+m2 = tm_shape(rnet_merged) + tm_lines("commute_fastest_bicycle_go_dutch", palette = "viridis", breaks = brks)
+tmap_arrange(m1, m2, nrow = 1, sync = TRUE)
+
 ```

### 2024-02-01 demo.qmd
fix typo

@@
 ```
-This is where rnet_merge funtion ard to handle the intersection road.
+This is where the rnet_merge function is hard to handle at the intersection road.
+
 ```{r}

### 2024-02-01 paper.qmd

---
title: "Approximate Network Matching"
author: 
  - name: Josiah Parry
    affiliation: Environmental Systems Research Institute, Redlands, CA, USA 
    orcid: 0000-0001-9910-865X
  - name: Robin Lovelace
    affiliation: Leeds Institute for Transport Studies, University of Leeds, UK
    orcid: 0000-0001-5679-6536
format: html
filters:
  - pseudocode
---

## Abstract

- A common issue in spatial data science is the reconciliation of two sets of linestrings. 
- linestrings may represent the same phenomenon but be topologically different
- joining data between these road networks is problematic for many reasons
- often there may be no join key present
  - if there is a join key, there remains an issue of wholesale joining of attributes
    - attributes of a linestring are intended be associated with that linestring, not another
    - we need to do a join and provide a weight for future calculations 
- in this paper we introduce a new algorithm to match two sets of linestrings
- it is implemented in an open source rust library with bindings to R and Python

## Problem Statement

TODO

## Existing algorithms

## Algorithm overview

The proposed algorithm aims to match elements of two sets of LineStrings that are topologically similar and estimate the amount of shared length between each pair of matched line strings.

Each LineString is composed of one or more Lines which is comprised of a single start or end point. The approximate network matching algorithm constructs two R* spatial indices over the component lines in $A$ and $B$. Intersection candidates between the two trees are used to limit the search space. For each candidate pair, the angle of the slopes are compared to determine if they are approximately parallel (parallelish). If the slopes are approximately parallel and the lines are within a minimum separable distance of each other, they are considered to match. The overlapping region between the matched lines is used to compute the shared length.

The result of the matching algorithm is a B-tree which can be used to generate a row-compressed sparse matrix.

### Identify match candidates

To identify matches between $A$ and $B$ we do not look at the LineStrings in their totality, but rather, by their individual components. $A$ and $B$ are comprised of one or more LineStrings index by $i$ and $j$ respectively. Each linestring is composed of one or more lines indexed as $k$. Matches are found between elements of $Aik$ and $Bjk$ using two R-trees.

We create an empty R-tree, $Tree_A$. For each line $Aik$ we compute the slope of the line and insert the geometry, slope, and index into the tree.

Next we create another empty R-tree, $Tree_B$, in which we will store each line in $Bjk$. However, instead of using the axis-aligned bounding box of $Bjk$, we create a newer, larger one, based on a distance tolerance, $DT$. The distance tolerance is used to expand the search for matches. We compute the AABB of $Bjk$, then expand the AABB by $DT$ in both the x and y directions. After doing so, we insert the geometry, slope, and index into $Tree_B$

If AABBs between $Tree_A$ and $Tree_B$ are intersecting, it means that that the lines $Aik$ and $Bjk$ might be within $DT$ of each other and should be checked to see if they are considered matches.

### Matching Criteria

Candidate matches as determined by intersecting AABBs must then be further evaluated.
Lines $Aik$ and $Bjk$ must be approximately parallel (parallelish) to be considered a match. To this end, an angle tolerance $AT$ is defined. We take the inverse tangent of the slopes of lines $Aik$ and $Bjk$ to find their angle. If the difference between these two angles are less than or equal to $AT$, we deem them tolerant or, parallelish.

Being confident that the $Aik$ and $Bjk$ are parallelish, we next need to determine if they are within the distance tolerance determined by $DT$. This is done by measuring the minimum separable distance between $Aik$ and $Bjk$. If both conditions are satisfied, then the lines are matched. Following, the shared segment length must be calculated.

### Caclulating segment overlap

Once two lines $Aik$ and $Bjk$ have been determined to be matches, we need to evaluate how much overlap exists between the two lines. This overlap is defined by the segment length of $Aik$ contained in the overlap in the x or y dimension between $Aik$ or $Bjk$.

Based on the angle of the line $A_{ik}$, $\theta_{A_{ik}}$, we either calculate the overlap in the line segments in either the x or y dimension.

![](assets/line-seg-overlap.png)

If $\theta_{Aik} \le 45^{\circ}$, we calculate the overlap between the range of x values of $A_{ik}$ and $B_{jk}$, $(x_{min}, x_{max})$. Using the slope of $A_{ik}$, solve for the values of y in the equation of the line. Using the calculated values of y, calculate the length of the line segment. If $\theta_{Aik} \gt 45^{\circ}$, we instead calculate the overlap in the range of y values and subsequently solve for x, then calculate the length of the line segment.

- Note that if there is no overlap in the x or y dimension and even if both matching criteria were met, there will be no shared length

## Algorithm Implementation

```pseudocode
#| label: alg-approx-net-matching
#| html-indent-size: "1.2em"
#| html-comment-delimiter: "//"
#| html-line-number: true
#| html-line-number-punc: ":"
#| html-no-end: false
#| pdf-placement: "htb!"
#| pdf-line-number: true

\begin{algorithm}
\caption{Approximate Network Matching}
\begin{algorithmic}
\State // Initialize R-trees for LineString components in sets A and B
\Procedure{ApproxNetworkMatch}{$A, B, DT, AT$}
  \State $Tree_A \gets$ InitializeEmptyRTree()
  \For{each $A_{ik} \in A$}
    \State $slope_{A_{ik}} \gets$ ComputeSlope($A_{ik}$)
    \State InsertIntoRTree($Tree_A, A_{ik}, slope_{A_{ik}}$)
  \EndFor

  \State $Tree_B \gets$ InitializeEmptyRTree()
  \For{each $B_{jk} \in B$}
    \State $expandedAABB_{B_{jk}} \gets$ ExpandAABB($B_{jk}, DT$)
    \State InsertIntoRTree($Tree_B, B_{jk}, expandedAABB_{B_{jk}}$)
  \EndFor

  \State // Identify potential match candidates
  \For{each pair $(A_{ik}, B_{jk})$ with intersecting AABBs}
    \If{IsParallelish($slope_{A_{ik}}, slope_{B_{jk}}, AT$) and IsWithinDistance($A_{ik}, B_{jk}, DT$)}
      \State // Calculate shared segment length
      \State $overlapLength \gets$ CalculateOverlapLength($A_{ik}, B_{jk}$)
      \State // Store matched pair and overlap length
      \State StoreMatchedPair($A_{ik}, B_{jk}, overlapLength$)
    \EndIf
  \EndFor

  \State \Return MatchedPairs
\EndProcedure

\State // Helper functions
\Function{IsParallelish}{$slope_{A}, slope_{B}, AT$}
  \State $angle_A \gets \arctan(slope_{A})$
  \State $angle_B \gets \arctan(slope_{B})$
  \State \Return $(|\angle_A - \angle_B| \le AT)$
\EndFunction

\Function{IsWithinDistance}{$A_{ik}, B_{jk}, DT$}
  \State $minDistance \gets$ ComputeMinSeparableDistance($A_{ik}, B_{jk}$)
  \State \Return $(minDistance \le DT)$
\EndFunction

\Function{CalculateOverlapLength}{$A_{ik}, B_{jk}$}
  \State $\theta_{A_{ik}} \gets$ ComputeAngle($A_{ik}$)
  \If{$\theta_{A_{ik}} \le 45^\circ$}
    \State $overlapLength \gets$ CalculateXOverlap($A_{ik}, B_{jk}$)
  \Else
    \State $overlapLength \gets$ CalculateYOverlap($A_{ik}, B_{jk}$)
  \EndIf
  \State \Return $overlapLength$
\EndFunction
\end{algorithmic}
\end{algorithm}

```

## Applications

- serves as the basis of joins and weighted aggregations
- NPT scottland 
- other possible use cases 
- can be used to subset networks based on shared characteristics e.g. road classification and matched between those subsets

## Discussion

- limitations: no 3d support, measurements in spherical coords are not as accurate

### 2024-02-01 r/demo.qmd
Move demo into r

--- 2024-02-01 demo.qmd
@@
 ---
+
+```{r}
+devtools::load_all()
+```
 
@@
 ```
+
 Here are the examples of an intersection road.
@@
 
-## With `{rsgeo}`
+## With Rust implementation
 

### 2024-02-01 r/demo.qmd
Update demo with reproducible code

@@
 library(sf)
+library(rsgeo)
 rnet_y = route_network_small["flow"]
@@
 ```
+
 This is where the rnet_merge function is hard to handle at the intersection road.
@@
 library(tmap)
-tmap_mode("view")
+tmap_mode("plot") # Set to view for interactive mode
 
@@
 
-Let's try with the `{rsgeo}` implementation:
-
 ```{r}
-library(rsgeo)
 rnet_x = rnet_x |>
@@
 ```{r}
-devtools::load_all()
 rnet_xp = rnet_x |>
@@
 system.time({
-rnet_matched = rnet_match(rnet_xp, rnet_yp, distance = 9, dist_chop = 1)
+rnet_matched = rnet_join_geos(rnet_xp, rnet_yp, distance = 9, dist_chop = 1)
 rnet_matched_agg = rnet_matched |>
@@
 ```{r}
-remotes::install_github("josiahparry/rsgeo", ref = "rent")
+# remotes::install_github("josiahparry/rsgeo", ref = "rent")
 

### 2024-02-01 r/demo.qmd
try geos on intersection road

@@
 ```{r}
-devtools::load_all()
+# devtools::load_all()
+# remotes::install_dev("rsgeo")
+# remotes::install_dev("stplanr")
 ```
@@
 library(stplanr)
+library(tmap)
 library(sf)
@@
 angle = 5
-rnet_merged = stplanr::rnet_merge(rnet_xp, rnet_yp, dist = dist, funs = funs, max_angle_diff = angle, segment_lentgh = 10)
+rnet_merged = stplanr::rnet_merge(rnet_xp, rnet_yp, dist = dist,  segment_length = 10, funs = funs, max_angle_diff = angle)
 
-library(tmap)
 tmap_mode("plot") # Set to view for interactive mode

### 2024-02-01 r/demo.qmd
adjust lines

@@
 ```
+```{}
+
+```
 
@@
 rnet_yp = sf::st_transform(rnet_intersection_complex, "EPSG:27700")
+
+rnet_xp <- line_segment(rnet_xp, segment_length = 10)
+rnet_yp <- line_segment(rnet_yp, segment_length = 10)
+
+result <- st_join(rnet_xp, rnet_yp, join = st_intersects)
+
 names(rnet_xp)
@@
 
-dist = 10
-angle = 5
-rnet_merged = stplanr::rnet_merge(rnet_xp, rnet_yp, dist = dist,  segment_length = 10, funs = funs, max_angle_diff = angle)
+dist = 20
+angle = 40
+rnet_merged = stplanr::rnet_merge(rnet_xp, rnet_yp, dist = dist, funs = funs, max_angle_diff = angle)
 
-tmap_mode("plot") # Set to view for interactive mode
+tmap_mode("view") # Set to view for interactive mode
 
@@
 ```{r}
-rnet_x_projected = st_transform(rnet_x, "EPSG:27700")
-rnet_y_projected = st_transform(rnet_y, "EPSG:27700")
+rnet_x = rnet_intersection_simple
+rnet_y = rnet_intersection_complex
+rnet_x_projected = st_transform(rnet_intersection_simple, "EPSG:27700")
+rnet_y_projected = st_transform(rnet_intersection_complex, "EPSG:27700")
 ```
@@
 )
-plot(rnet_xlbcu, col = "lightgrey")
+plot(rnet_xlbcu, col = "grey" )
+plot(rnet_y_geos, col = "blue", add = TRUE)
+plot(rnet_y_remove, col = "red",add = TRUE)
 ```
@@
 )
-plot(rnet_ycj)
+plot(rnet_ycl)
 ```
@@
 rnet_x_df_expanded = rnet_x_df[rnet_ycj$y, ]
-rnet_y_df_expanded$osm_id = rnet_x_df_expanded$osm_id
+
+rnet_y_df_expanded$index = rnet_x_df_expanded$index
 nrow(rnet_y_df_expanded)
-nrow(rnet_x)
+nrow(rnet_y)
 rnet_x_agg = rnet_y_df_expanded |>
-  dplyr::group_by(osm_id) |>
+  dplyr::group_by(index) |>
   dplyr::summarise(
-    flow = sum(flow),
-    quietness = mean(quietness)
+    commute_fastest_bicycle_go_dutch = sum(commute_fastest_bicycle_go_dutch),
+    quietness = mean(Quietness)
   )
@@
 )
-plot(rnet_x_joined["flow"])
+
+tmap_mode("view") # Set to view for interactive mode
+
+brks = c(0, 50, 100, 200, 500,1000,2000)
+m1 = tm_shape(rnet_x_joined) + tm_lines("commute_fastest_bicycle_go_dutch", palette = "viridis", breaks = brks)
+m2 = tm_shape(rnet_y) + tm_lines("commute_fastest_bicycle_go_dutch", palette = "viridis", breaks = brks)
+tmap_arrange(m1, m2, nrow = 1, sync = TRUE)
+
 ```
@@
 plot(rnet_y$geometry, lwd = 5, col = "lightgrey")
-plot(rnet_x_joined["flow"], add = TRUE, lwd = 2)
+plot(rnet_x_joined["commute_fastest_bicycle_go_dutch"], add = TRUE, lwd = 2)
 par(mfrow = c(1, 1))

### 2024-02-01 r/demo.qmd
Add more examples

@@
 ```{r}
-# devtools::load_all()
+devtools::load_all()
 # remotes::install_dev("rsgeo")
@@
 plot(rnet_x_e, add = TRUE, pch = 3)
-```
-```{}
-
 ```
@@
 
-rnet_xp <- line_segment(rnet_xp, segment_length = 10)
-rnet_yp <- line_segment(rnet_yp, segment_length = 10)
+# rnet_xp <- line_segment(rnet_xp, segment_length = 10)
+# rnet_yp <- line_segment(rnet_yp, segment_length = 10)
 
@@
 
-tmap_mode("view") # Set to view for interactive mode
+tmap_mode("plot") # Set to view for interactive mode
 
 brks = c(0, 50, 100, 200, 500,1000,2000)
-m1 = tm_shape(rnet_intersection_complex) + tm_lines("commute_fastest_bicycle_go_dutch", palette = "viridis", breaks = brks)
-m2 = tm_shape(rnet_merged) + tm_lines("commute_fastest_bicycle_go_dutch", palette = "viridis", breaks = brks)
+m1 = tm_shape(rnet_intersection_complex) + tm_lines("commute_fastest_bicycle_go_dutch", palette = "viridis", breaks = brks, lwd = 5)
+m2 = tm_shape(rnet_merged) + tm_lines("commute_fastest_bicycle_go_dutch", palette = "viridis", breaks = brks, lwd = 5)
 tmap_arrange(m1, m2, nrow = 1, sync = TRUE)
@@
 
-## With `{geos}`
+## `rnet_join_geos`
+
+```{r}
+rnet_xp = rnet_xp |> 
+  dplyr::mutate(identifier = 1:nrow(rnet_xp)) |> 
+  dplyr::select(identifier, dplyr::everything())
+rnet_yp = rnet_yp |> 
+  dplyr::transmute(value = commute_fastest_bicycle_go_dutch)
+rnet_matched = rnet_join_geos(rnet_xp, rnet_yp, distance = 9, dist_chop = 1)
+rnet_matched_agg = rnet_matched |>
+  dplyr::group_by(identifier) |>
+  dplyr::summarise(value = sum(value))
+rnet_joined = rnet_xp |>
+  dplyr::left_join(rnet_matched_agg)
+summary(rnet_joined)
+# rnet_joined$value
+plot(rnet_joined["value"], lwd = 3)
+plot(rnet_yp["value"], lwd = 3)
+plot(rnet_matched)
+```
+
+## With `{geos}`: details
 
@@
 
-tmap_mode("view") # Set to view for interactive mode
+tmap_mode("plot") # Set to view for interactive mode
 
@@
 ```{r}
+#| eval: false
 rnet_x = rnet_x_projected
@@
 ```{r}
+#| eval: false
 rnet_x = sf::read_sf("https://github.com/ropensci/stplanr/releases/download/v1.0.2/rnet_x_ed.geojson")
@@
 ```{r}
+#| eval: false
 plot(rnet_y["value"], lwd = 5)
@@
 ```{r}
+#| eval: false
 plot(rnet_x$geometry, lwd = 9, col = "lightgrey")
@@
 ```{r}
+#| eval: false
 # Stplanr implementation:
@@
 ```{r}
+#| eval: false
 rnet_xp = rnet_x |>
@@
 ```{r}
+#| eval: false
 # remotes::install_github("josiahparry/rsgeo", ref = "rent")

### 2024-02-01 r/R/rnet_match_geos.R

@@ -43,7 +43,9 @@ rnet_match_geos = function(
     rnet_x_geos,
     rnet_y_geos,
     distance = 9,
-    dist_chop = 0.1) {
+    dist_chop = 0.1,
+    predicate = "within"
+    ) {
   params = geos::geos_buffer_params(end_cap_style = "flat")
   rnet_x_buffer = geos::geos_buffer(rnet_x_geos, distance, params = params)
   rnet_xbl = geos::geos_boundary(rnet_x_buffer)
@@ -56,7 +58,8 @@ rnet_match_geos = function(
   )
   rnet_ycj = geos::geos_inner_join_keys(
     rnet_y_chopped,
-    rnet_x_buffer
+    rnet_x_buffer,
+    predicate = predicate
   )
   rnet_ycj
 }

### 2024-02-03 r/R/aggregate.R
draft rnet_aggregate()

@@ -1,0 +1,51 @@
+#' Only works with numeric variables
+rnet_aggregate <- function(
+    x, y, matches, ...,
+    # we automatically calculate the length of y
+    # if sf::st_length() doesn't work it must be supplied
+    y_len = sf::st_length(y)
+) {
+  # TODO object validation of x, y, and matches
+
+  # capture variables
+  vars <- rlang::ensyms(...)
+  # get var-names
+  var_names <- vapply(vars, rlang::as_string, character(1))
+  # TODO validate variables are in y before subsetting
+  # extract j index
+  j <- matches$j
+  # subset vars by j to get ij pairs
+  ij <- rlang::set_names(lapply(var_names, \(.x) y[[.x]][j]), var_names)
+  # combine into 1 df
+  dplyr::bind_cols(matches, ij) |>
+    dplyr::mutate(
+      wt = shared_len / y_len[j]
+    ) |>
+    dplyr::group_by(i) |>
+    dplyr::summarise(dplyr::across(
+      -all_of(c("j", "shared_len", "wt")),
+      ~sum(.x * wt, na.rm = TRUE)
+    ))
+}
+
+
+# library(sf)
+# library(dplyr)
+#
+# x <- read_sf("data-raw/geojson/intersection_example_simple.geojson") |>
+#   sf::st_transform(27700)
+# y <- read_sf("data-raw/geojson/intersection_example_complex.geojson") |>
+#   sf::st_transform(27700)
+#
+# matches <- rnetmatch::rnet_match(x, y, 10, 5) |>
+#   as_tibble()
+#
+# rnet_aggregate(
+#   x, y, matches,
+#   # columns in j
+#   all_fastest_bicycle,
+#   Quietness,
+#   commute_quietest_bicycle_go_dutch
+# )
+#
+#

### 2024-02-03 r/R/aggregate.R
update doc

@@ -1,4 +1,11 @@
-#' Only works with numeric variables
+#' Weighted aggregation of matched networks
+#'
+#' @param x data.frame
+#' @param y sf object
+#' @param matches result of rnet_match()
+#' @param ... unquoted variable names in y
+#' @param y_len default `sf::st_length(y)` a numeric vector the same length as y containing the length of each linestring in y
+#' @export
 rnet_aggregate <- function(
     x, y, matches, ...,
     # we automatically calculate the length of y

### 2024-02-04 r/demo.qmd
Add length travelled and unevaluated python demo

@@
 rnet_merged = stplanr::rnet_merge(rnet_xp, rnet_yp, dist = dist, funs = funs, max_angle_diff = angle)
-
+# Explore distance travelled:
+sum(rnet_merged$commute_fastest_bicycle_go_dutch * sf::st_length(rnet_merged$geometry))
+# And for the complex network:
+sum(rnet_yp$commute_fastest_bicycle_go_dutch * sf::st_length(rnet_yp$geometry))
 tmap_mode("plot") # Set to view for interactive mode
@@
 ```
+
+<!-- # In Python -->
+
+```{python}
+#| eval: false
+#| echo: false
+import geopandas as gpd
+y = gpd.read_file("https://github.com/nptscot/rnetmatch/releases/download/intersection_road/intersection_example_complex.geojson")
+y.plot()
+# show it:
+import matplotlib.pyplot as plt
+plt.show()
+
+x = gpd.read_file("https://github.com/nptscot/rnetmatch/releases/download/intersection_road/intersection_example_simple.geojson")
+x.plot()
+```

### 2024-02-04 paper.qmd

@@
 
+```{r, include = FALSE}
+knitr::opts_chunk$set(echo = FALSE)
+```
+
 ## Abstract
+
+Reconciling topologically different linestrings is a fundamental challenge in spatial data science, particularly when attempting to integrate network datasets from disparate sources.
+Existing methods often struggle with the absence of join keys and the need for wholesale joining of attributes, hindering effective data integration and analysis.
+In this paper, we propose a novel algorithm for matching two sets of linestrings, implemented in an open-source Rust library with bindings to R and Python.
+Our algorithm addresses the challenge by identifying topologically similar linestrings and estimating the shared length between each pair of matched linestrings.
+By leveraging R* spatial indices and angle-based matching criteria, our approach effectively reconciles linestrings with varying topologies.
+We demonstrate the utility of our algorithm through applications in spatial data analysis, including joins, weighted aggregations, and network subset identification based on shared characteristics.
+The proposed algorithm offers a robust solution for reconciling linestrings in spatial datasets, with implications for various domains in spatial data science.
+
+<!--
 
@@
 - it is implemented in an open source rust library with bindings to R and Python
+-->
 
@@
 Each linestring is composed of one or more lines indexed as $k$.
-Matches are found between elements of $Aik$ and $Bjk$ using two R-trees.
+Matches are found between elements of $A_{ik}$ and $B_{jk}$ using two R-trees.
 
-We create an empty R-tree, $Tree_A$. For each line $Aik$ we compute the slope of the line and insert the geometry, slope, and index into the tree.
+We create an empty R-tree, $Tree_A$. For each line $A_{ik}$ we compute the slope of the line and insert the geometry, slope, and index into the tree.
 
-Next we create another empty R-tree, $Tree_B$, in which we will store each line in $Bjk$.
-However, instead of using the axis-aligned bounding box of $Bjk$, we create a newer, larger one, based on a distance tolerance, $DT$.
+Next we create another empty R-tree, $Tree_B$, in which we will store each line in $B_{jk}$.
+However, instead of using the axis-aligned bounding box of $B_{jk}$, we create a newer, larger one, based on a distance tolerance, $DT$.
 The distance tolerance is used to expand the search for matches.
-We compute the AABB of $Bjk$, then expand the AABB by $DT$ in both the x and y directions.
+We compute the AABB of $B_{jk}$, then expand the AABB by $DT$ in both the x and y directions.
 After doing so, we insert the geometry, slope, and index into $Tree_B$
 
-If AABBs between $Tree_A$ and $Tree_B$ are intersecting, it means that that the lines $Aik$ and $Bjk$ might be within $DT$ of each other and should be checked to see if they are considered matches.
+```{r message = FALSE, echo = FALSE}
+#| layout-ncol: 2
+suppressMessages(library(sf))
+library(rsgeo)
+library(ggplot2)
+library(patchwork)
+conflicted::conflict_prefer("ggplot2", "rsgeo")
+
+# box to crop geometry to
+crop_box <- st_bbox(c("xmin" = 427200, xmax = 427500, ymin = 433550, ymax = 433700))
+
+rnet_x <- "https://raw.githubusercontent.com/nptscot/networkmerge/main/data/rnet_armley.geojson" |> 
+  read_sf() |>
+  st_geometry() |> 
+  st_transform(27700) |> 
+  st_crop(crop_box)
+
+rnet_y <- "https://raw.githubusercontent.com/nptscot/networkmerge/main/data/rnet_armley_line.geojson" |>
+  read_sf() |> 
+  st_crop(crop_box)
+
+x <- as_rsgeo(sf::st_transform(rnet_x, 27700))
+y <- as_rsgeo(st_transform(rnet_y, 27700))
+
+# axis-aligned-bounding-box for x
+xbb <- bounding_rect(explode_lines(x))
+
+# creating bounding rects for y
+# need to expand them
+ybb <- bounding_rect(explode_lines(y))
+
+# define function to expand the AABB
+expand_aabb <- function(x, DT) {
+  crds <- coords(x)
+  # xmin, max, max, min, min
+  # ymin min max max min
+  crds[,1] <- crds[,1] + (c(-1, 1, 1, -1, -1)  * DT)
+  crds[,2] <- crds[,2] + (c(-1, -1, 1, 1, -1)  * DT)
+  rsgeo::geom_polygon(crds$x, crds$y, crds$polygon_id)
+}
+
+DT <- 2.5
+xbb_sf <- st_as_sfc(xbb) |> st_set_crs(27700)
+ybb_sf <- st_as_sfc(expand_aabb(ybb, DT)) |> st_set_crs(27700)
+
+p1 <- ggplot() +
+  geom_sf(
+    data = xbb_sf, 
+    fill = "#76b5c5", alpha = 0.25, lwd = 0.1
+  ) +
+  geom_sf(data = rnet_x, lwd = 0.2) +
+  labs(title = "Axis-aligned-bounding-boxes of A") + 
+  theme_void()
+
+p2 <- ggplot() +
+  geom_sf(
+    data = ybb_sf,
+    fill = "#e28743", alpha = 0.25,lwd = 0.1
+  ) +
+  geom_sf(data = rnet_y, alpha = 0.5) +
+  labs(
+    title = "Axis-aligned-bounding-boxes of B",
+    subtitle = paste0(DT, "meter distance tolerance")
+  ) + 
+  theme_void()
+
+p3 <- ggplot() +
+  geom_sf(
+    data = xbb_sf, 
+    fill = "#76b5c5", alpha = 0.25, lwd = 0.1
+  ) + 
+  geom_sf(
+    data = ybb_sf,
+    fill = "#e28743", alpha = 0.25, lwd = 0.1
+  ) + 
+  geom_sf(data = rnet_x, alpha = 0.5, lwd = 0.15) +
+  geom_sf(data = rnet_y, alpha = 0.5) +
+  labs(title = "AABB and Networks Overlay") + 
+  theme_void()
+
+p1
+p2
+```
+
+If AABBs between $Tree_A$ and $Tree_B$ are intersecting, it means that that the lines $A_{ik}$ and $B_{jk}$ might be within $DT$ of each other and should be checked to see if they are considered matches.
+
+```{r}
+p3
+```
 
@@
 Candidate matches as determined by intersecting AABBs must then be further evaluated.
-Lines $Aik$ and $Bjk$ must be approximately parallel (parallelish) to be considered a match.
+Lines $A_{ik}$ and $B_{jk}$ must be approximately parallel (parallelish) to be considered a match.
 To this end, an angle tolerance $AT$ is defined.
-We take the inverse tangent of the slopes of lines $Aik$ and $Bjk$ to find their angle.
+We take the inverse tangent of the slopes of lines $A_{ik}$ and $B_{jk}$ to find their angle.
 If the difference between these two angles are less than or equal to $AT$, we deem them tolerant or, parallelish.
 
-Being confident that the $Aik$ and $Bjk$ are parallelish, we next need to determine if they are within the distance tolerance determined by $DT$.
-This is done by measuring the minimum separable distance between $Aik$ and $Bjk$.
+```{r}
+#| fig-cap: "Matched lines with 15° angle tolerance and 2.5 meter distance tolerance."
+mtx <- rnetmatch::rnet_match(rnet_x, rnet_y, 2.5, 15)
+
+i <- 4
+ydx <- which(mtx$i == i)
+xx <- rnet_x[i]
+yy <- rnet_y$geometry[mtx$j[ydx]]
+
+# plot the lines that we will measure 
+plot(xx, lty = 3)
+plot(yy, add = TRUE)
+```
+
+Being confident that the $A_{ik}$ and $B_{jk}$ are parallelish, we next need to determine if they are within the distance tolerance determined by $DT$.
+This is done by measuring the minimum separable distance between $A_{ik}$ and $B_{jk}$.
 If both conditions are satisfied, then the lines are matched.
@@
 
-Once two lines $Aik$ and $Bjk$ have been determined to be matches, we need to evaluate how much overlap exists between the two lines.
-This overlap is defined by the segment length of $Aik$ contained in the overlap in the x or y dimension between $Aik$ or $Bjk$.
+Once two lines $A_{ik}$ and $B_{jk}$ have been determined to be matches, we need to evaluate how much overlap exists between the two lines.
+This overlap is defined by the segment length of $A_{ik}$ contained in the overlap in the x or y dimension between $A_{ik}$ or $B_{jk}$.
 
@@
 
-![](assets/line-seg-overlap.png)
+![](assets/line-seg-overlap-top.png)
 
-If $\theta_{Aik} \le 45^{\circ}$, we calculate the overlap between the range of x values of $A_{ik}$ and $B_{jk}$, $(x_{min}, x_{max})$.
+If $\theta_{A_{ik}} \le 45^{\circ}$, we calculate the overlap between the range of x values of $A_{ik}$ and $B_{jk}$, $(x_{min}, x_{max})$.
 Using the slope of $A_{ik}$, solve for the values of y in the equation of the line.
 Using the calculated values of y, calculate the length of the line segment.
-If $\theta_{Aik} \gt 45^{\circ}$, we instead calculate the overlap in the range of y values and subsequently solve for x, then calculate the length of the line segment.
-
-- Note that if there is no overlap in the x or y dimension and even if both matching criteria were met, there will be no shared length
+If $\theta_{A_{ik}} \gt 45^{\circ}$, we instead calculate the overlap in the range of y values and subsequently solve for x, then calculate the length of the line segment.
+Note that if there is no overlap in the x or y dimension and even if both matching criteria were met, there will be no shared length, and the resultant value with be 0.
 
@@
     \State $slope_{A_{ik}} \gets$ ComputeSlope($A_{ik}$)
-    \State InsertIntoRTree($Tree_A, A_{ik}, slope_{A_{ik}}$)
+    \State InsertIntoRTree($Tree_A, i, A_{ik}, slope_{A_{ik}}$)
   \EndFor
@@
     \State $expandedAABB_{B_{jk}} \gets$ ExpandAABB($B_{jk}, DT$)
-    \State InsertIntoRTree($Tree_B, B_{jk}, expandedAABB_{B_{jk}}$)
+    \State InsertIntoRTree($Tree_B, j, B_{jk}, expandedAABB_{B_{jk}}$)
   \EndFor
@@
       \State // Store matched pair and overlap length
-      \State StoreMatchedPair($A_{ik}, B_{jk}, overlapLength$)
+      \State StoreOrUpdateMatchedPair($A_{ik}, B_{jk}, overlapLength$)
     \EndIf
@@
   \State $angle_B \gets \arctan(slope_{B})$
-  \State \Return $(|\angle_A - \angle_B| \le AT)$
+  \State \Return $(|angle_A - angle_B| \le AT)$
 \EndFunction
@@
 
-- limitations: no 3d support, measurements in spherical coords are not as accurate
+This algorithm is limited in that it is designed in $\Bbb{R}^2$ space.
+As such, there is no support for 3-dimensional or spherical coordinates.
+It is, however, conceivable to apply the same principles to these scenarios.
+The challenge, then, is to determine overlap regions, lengths, and lines-segments.

### 2024-02-04 r/R/aggregate.R
Tweak aggregation code

@@ -6,12 +6,24 @@
 #' @param ... unquoted variable names in y
 #' @param y_len default `sf::st_length(y)` a numeric vector the same length as y containing the length of each linestring in y
 #' @export
+#' @examples
+#' library(sf)
+#' x <- read_sf("data-raw/geojson/princes_street_minimal_x_1.geojson") |>
+#'   sf::st_transform(27700)
+#' y <- read_sf("data-raw/geojson/princes_street_minimal.geojson") |>
+#'   sf::st_transform(27700)
+#' matches <- rnetmatch::rnet_match(x, y, dist_tolerance = 10, angle_tolerance = 5)
+#' y_aggregated <- rnet_aggregate(x, y, matches, value)
+#' y_joined <- cbind(x, y_aggregated)
+#' plot(y["value"], lwd = 5)
+#' plot(y_joined["value"], lwd = 5)
+#' y_y_joined <- dplyr::bind_rows(y["value"], y_joined["value"])
+#' plot(y_y_joined, lwd = 5)
 rnet_aggregate <- function(
     x, y, matches, ...,
     # we automatically calculate the length of y
     # if sf::st_length() doesn't work it must be supplied
-    y_len = sf::st_length(y)
-) {
+    y_len = as.numeric(sf::st_length(y))) {
   # TODO object validation of x, y, and matches
 
   # capture variables
@@ -31,7 +43,7 @@ rnet_aggregate <- function(
     dplyr::group_by(i) |>
     dplyr::summarise(dplyr::across(
       -all_of(c("j", "shared_len", "wt")),
-      ~sum(.x * wt, na.rm = TRUE)
+      ~ sum(.x * wt, na.rm = TRUE)
     ))
 }
 

### 2024-02-04 r/R/aggregate.R
Fix typo

@@ -43,7 +43,7 @@ rnet_aggregate <- function(
     dplyr::group_by(i) |>
     dplyr::summarise(dplyr::across(
       -all_of(c("j", "shared_len", "wt")),
-      ~ sum(.x * wt, na.rm = TRUE)
+      ~sum(.x * wt, na.rm = TRUE)
     ))
 }
 

### 2024-02-05 r/demo.qmd

@@
 
-This is where the rnet_merge function is hard to handle at the intersection road.
+## Minimal example
+
+A minimal example can be seen below:
+
+```{r}
+# setup input dataset:
+x = sf::read_sf("data-raw/geojson/princes_street_minimal_x_1.geojson") |>
+  sf::st_transform(27700)
+y = sf::read_sf("data-raw/geojson/princes_street_minimal.geojson") |>
+  sf::st_transform(27700)
+
+x = x[3, ]
+y = y[2:3, ]
+matches <- rnetmatch::rnet_match(x, y, dist_tolerance = 10, angle_tolerance = 5)
+y_aggregated <- rnet_aggregate(x, y, matches, value)
+y_joined <- cbind(x, y_aggregated)
+y_joined
+
+x_segmented = stplanr::line_segment(x, segment_length = 10)
+matches <- rnetmatch::rnet_match(x_segmented, y, dist_tolerance = 10, angle_tolerance = 5)
+y_joined_segmented <- rnet_aggregate(x_segmented, y, matches, value)
+sum(y_joined_segmented$value) # 
+```
+
+```{r}
+funs = list(value = sum)
+# With stplanr:
+rnet_merged = stplanr::rnet_merge(x, y, dist = 10, funs = funs, max_angle_diff = 5, sum_flows = FALSE)
+rnet_merged_segmented = stplanr::rnet_merge(x_segmented, y, dist = 10, funs = funs, max_angle_diff = 5, sum_flows = FALSE)
+rnet_merged$value
+rnet_merged_segmented$value
+```
+
+## Intersection example
+
+This is where the rnet_merge function struggles to handle the intersection road.
 

### 2024-02-13 r/R/aggregate.R
draft aggregate functions for intensive and extensive

@@ -6,7 +6,8 @@
 #' @param ... unquoted variable names in y
 #' @param y_len default `sf::st_length(y)` a numeric vector the same length as y containing the length of each linestring in y
 #' @export
-rnet_aggregate <- function(
+#' @rdname aggregate
+rnet_aggregate_extensive <- function(
     x, y, matches, ...,
     # we automatically calculate the length of y
     # if sf::st_length() doesn't work it must be supplied
@@ -26,7 +27,7 @@ rnet_aggregate <- function(
   # combine into 1 df
   dplyr::bind_cols(matches, ij) |>
     dplyr::mutate(
-      wt = shared_len / y_len[j]
+      wt = shared_len / as.numeric(y_len[j])
     ) |>
     dplyr::group_by(i) |>
     dplyr::summarise(dplyr::across(
@@ -36,23 +37,35 @@ rnet_aggregate <- function(
 }
 
 
-# library(sf)
-# library(dplyr)
-#
-# x <- read_sf("data-raw/geojson/intersection_example_simple.geojson") |>
-#   sf::st_transform(27700)
-# y <- read_sf("data-raw/geojson/intersection_example_complex.geojson") |>
-#   sf::st_transform(27700)
-#
-# matches <- rnetmatch::rnet_match(x, y, 10, 5) |>
-#   as_tibble()
-#
-# rnet_aggregate(
-#   x, y, matches,
-#   # columns in j
-#   all_fastest_bicycle,
-#   Quietness,
-#   commute_quietest_bicycle_go_dutch
-# )
-#
-#
+#' @export
+#' @rdname aggregate
+rnet_aggregate_intensive <- function(
+    x, y, matches, ...,
+    # we automatically calculate the length of y
+    # if sf::st_length() doesn't work it must be supplied
+    x_len = sf::st_length(x)
+) {
+  # TODO object validation of x, y, and matches
+
+  # capture variables
+  vars <- rlang::ensyms(...)
+  # get var-names
+  var_names <- vapply(vars, rlang::as_string, character(1))
+  # TODO validate variables are in y before subsetting
+  # extract j index
+  j <- matches$j
+  # subset vars by j to get ij pairs
+  ij <- rlang::set_names(lapply(var_names, \(.x) y[[.x]][j]), var_names)
+  # combine into 1 df
+  dplyr::bind_cols(matches, ij) |>
+    dplyr::mutate(
[…]

### 2024-02-13 r/R/aggregate.R
update with example and merge conflict resolution

@@ -8,22 +8,22 @@
 #' @export
 #' @examples
 #' library(sf)
-#' x <- read_sf("data-raw/geojson/princes_street_minimal_x_1.geojson") |>
-#'   sf::st_transform(27700)
-#' y <- read_sf("data-raw/geojson/princes_street_minimal.geojson") |>
-#'   sf::st_transform(27700)
-#' matches <- rnetmatch::rnet_match(x, y, dist_tolerance = 10, angle_tolerance = 5)
-#' y_aggregated <- rnet_aggregate(x, y, matches, value)
-#' y_joined <- cbind(x, y_aggregated)
-#' plot(y["value"], lwd = 5)
-#' plot(y_joined["value"], lwd = 5)
-#' y_y_joined <- dplyr::bind_rows(y["value"], y_joined["value"])
-#' plot(y_y_joined, lwd = 5)
-rnet_aggregate <- function(
+#' xfp <- system.file("extdata/princes_street_minimal_x_1.geojson", package = "rnetmatch")
+#' yfp <- system.file("extdata/princes_street_minimal.geojson", package = "rnetmatch")
+#'
+#' x <- st_transform(st_read(xfp, quiet = TRUE), 27700)[3, ]
+#' y <- st_transform(st_read(yfp, quiet = TRUE), 27700)[2:3, ]
+#'
+#' matches <- rnet_match(x, y, dist_tolerance = 10, angle_tolerance = 5)
+#' rnet_aggregate_intensive(x, y, matches, value)
+#' rnet_aggregate_extensive(x, y, matches, value)
+#' @rdname aggregate
+rnet_aggregate_extensive <- function(
     x, y, matches, ...,
     # we automatically calculate the length of y
     # if sf::st_length() doesn't work it must be supplied
-    y_len = as.numeric(sf::st_length(y))) {
+    y_len = as.numeric(sf::st_length(y))
+  ) {
   # TODO object validation of x, y, and matches
 
   # capture variables
@@ -38,33 +38,43 @@ rnet_aggregate <- function(
   # combine into 1 df
   dplyr::bind_cols(matches, ij) |>
     dplyr::mutate(
-      wt = shared_len / y_len[j]
+      wt = shared_len / as.numeric(y_len[j])
     ) |>
     dplyr::group_by(i) |>
     dplyr::summarise(dplyr::across(
       -all_of(c("j", "shared_len", "wt")),
-      ~sum(.x * wt, na.rm = TRUE)
[…]

### 2024-02-13 r/demo.qmd
Update with more demos

@@
 y_joined_segmented <- rnet_aggregate(x_segmented, y, matches, value)
-sum(y_joined_segmented$value) # 
+y_joined_segmented
+sum(y_joined_segmented$value)
 ```

### 2024-02-13 r/demo.qmd
More demos

--- 2024-02-04 r/demo.qmd
@@
 plot(rnet_x_e, add = TRUE, pch = 3)
+```
+
+We can merge the two datasets using the `rnet_merge` function from the `stplanr` package:
+
+```{r}
+rnet_merged = rnet_merge(rnet_x, rnet_y["flow"], dist = 9, segment_length = 20, funs = list(flow = sum))
+sum(rnet_merged$flow * sf::st_length(rnet_merged$geometry), na.rm = TRUE)
+sum(rnet_y$flow * sf::st_length(rnet_y$geometry), na.rm = TRUE)
+```
+
+We can tweak the parameters:
+
+```{r}
+rnet_merged = rnet_merge(rnet_x, rnet_y["flow"], dist = 9, segment_length = 20, funs = list(flow = sum), sum_flows = FALSE)
+sum(rnet_merged$flow * sf::st_length(rnet_merged$geometry), na.rm = TRUE)
+```
+
+With the new implementation:
+
+```{r}
+
+```
+
+# With the Princes Street example
+
+```{r}
+library(sf)
+library(rnetmatch)
+x <- read_sf(system.file("extdata", "princes_street_minimal_x_1.geojson", package = "rnetmatch")) |>
+  sf::st_transform(27700)
+x = x[3, ]
+x
+y <- read_sf(system.file("extdata", "princes_street_minimal.geojson", package = "rnetmatch")) |>
+  sf::st_transform(27700)
+y = y[2:3, ]
+
+plot(x$geometry, col = "grey", lwd = 5)
+plot(y$geometry, add = TRUE, lwd = 2)
+
+matches <- rnetmatch::rnet_match(x, y, dist_tolerance = 10, angle_tolerance = 5)
+
+rnet_aggregate_intensive(x, y, matches, value)
+rnet_aggregate_extensive(x, y, matches, value)
+```
+
+For the full example:
+
+```{r}
+x <- read_sf(system.file("extdata", "princes_street_minimal_x_1.geojson", package = "rnetmatch")) |>
+  sf::st_transform(27700)
+x
+y <- read_sf(system.file("extdata", "princes_street_minimal.geojson", package = "rnetmatch")) |>
+  sf::st_transform(27700)
+
+my = tm_shape(y) + tm_lines("value", lwd = 2)
+mx = tm_shape(x) + tm_lines(lwd = 5, col = "grey")
+mx + my
+
+matches <- rnetmatch::rnet_match(x, y, dist_tolerance = 10, angle_tolerance = 5)
+
+mx_intensive = rnet_aggregate_intensive(x, y, matches, value)
+mx_extensive = rnet_aggregate_extensive(x, y, matches, value)
+mx_intensive
+mx_extensive
+mx_intensivej = dplyr::left_join(x, mx_intensive |> dplyr::rename(id = i))
+mx_extensivej = dplyr::left_join(x, mx_extensive |> dplyr::rename(id = i))
+
+my + tm_shape(mx_intensivej) + tm_lines("value", lwd = 2)
+my + tm_shape(mx_extensivej) + tm_lines("value", lwd = 2)
 ```

### 2024-02-14 r/demo.qmd
Update demo

@@
 
-Here are the examples of an intersection road.
+## Minimal example
+
+A minimal example can be seen below:
+
+```{r}
+# setup input dataset:
+x = sf::read_sf("data-raw/geojson/princes_street_minimal_x_1.geojson") |>
+  sf::st_transform(27700)
+y = sf::read_sf("data-raw/geojson/princes_street_minimal.geojson") |>
+  sf::st_transform(27700)
+x = x[3, ]
+y = y[2:3, ]
+matches <- rnetmatch::rnet_match(x, y, dist_tolerance = 10, angle_tolerance = 5)
+y_aggregated <- rnet_aggregate(x, y, matches, value)
+y_joined <- cbind(x, y_aggregated)
+y_joined
+x_segmented = stplanr::line_segment(x, segment_length = 10)
+matches <- rnetmatch::rnet_match(x_segmented, y, dist_tolerance = 10, angle_tolerance = 5)
+y_joined_segmented <- rnet_aggregate(x_segmented, y, matches, value)
+sum(y_joined_segmented$value) # 
+```
+
+```{r}
+funs = list(value = sum)
+# With stplanr:
+rnet_merged = stplanr::rnet_merge(x, y, dist = 10, funs = funs, max_angle_diff = 5, sum_flows = FALSE)
+rnet_merged_segmented = stplanr::rnet_merge(x_segmented, y, dist = 10, funs = funs, max_angle_diff = 5, sum_flows = FALSE)
+rnet_merged$value
+rnet_merged_segmented$value
+```
+
+## Intersection example
+
+This is where the rnet_merge function struggles to handle the intersection road.
 

### 2024-02-18 paper.qmd
miscellaneous updates

@@
   - pseudocode
+bibliography: refs.bib
 ---
@@
 
-TODO
+**Overarching Idea**:
+
+- use the approximate network matching to create a list of _very strong_ candidate matches with a known shared overlap amount. 
+- The result of the approximate network match can be taken and utilized for vertical conflation using many different approaches 
+- in this paper we introduce a method of spatial interpolation of linear features
+
+> "Due to the complexity and limitations of existing methods, planners and analysts often have to employ a heavily manual conflation process, which is time‐consuming and often prohibitively expensive." Lei, Ting; Lei, Zhen (2019).
 
 ## Existing algorithms
+
+Goodchild and Hunter 1997 introduce a buffer based algorithm which calculates the proportion of length covered by a target and source feature. They state that
+
+> we suggest that a similar approach based on the percentage of the reference source length rather than the tested source might be more related to generalization...
+
+- Our algorithm, due to its reporting of shared overlap permits us to adapt the Goodchild algorithm and use % source length overlap as a cut off. 
+  - The issue here is that there may be complete overlap from the source but the source may continue well past. we must address this. This is why we will lean on areal weighted interpolation
+
+- (Chehreghan & Abbaspour, 2018) use a genetic algorithm. These are computational expensive and do not sacle well
+
+- overline approach
+
+### Spatial Criteria
+
+our algorithm can be thought of as an extension of buffer and orientation analysis
+
+- location / proximity measured as a distance buffer (of sorts)
+- orientation (angle)
+- length
 
@@
 
-## Applications
+## Interpolation of values
 
-- serves as the basis of joins and weighted aggregations
-- NPT scottland 
-- other possible use cases 
-- can be used to subset networks based on shared characteristics e.g. road classification and matched between those subsets
+<!-- ## Applications  -->
+
+<!-- - serves as the basis of joins and weighted aggregations -->
+<!-- - NPT scottland  -->
+<!-- - other possible use cases  -->
+<!-- - can be used to subset networks based on shared characteristics e.g. road classification and matched between those subsets  -->
 
@@
 
-This algorithm is limited in that it is designed in $\Bbb{R}^2$ space.
+- This algorithm is limited in that it is designed in $\Bbb{R}^2$ space.
 As such, there is no support for 3-dimensional or spherical coordinates.
@@
 The challenge, then, is to determine overlap regions, lengths, and lines-segments.
+
+- 3D could be important though not many network datasets keep track of the height of points. Otherwise, measuring distance in 3 dimension for candidates could be useful.
+
+- limited in that the geometries must be in the same coordinate system and must be proximal to eachother. Does not work for arbitrarily scaled geometries
+
+- does not handle geometry shift (Lei, Ting; Lei, Zhen (2019))
+  - we make no affine transformations to handle shifts in geometries
+
+------
+
+## Notes
+
+- conflation between two datasets that represent the same area is called "vertical conflation" - JCS technical report
+
+- "version matching" (JCS) features are assumed to be the same and transferring of attributes
+
+- This is a problem of geometric conflation, trying to incorporate attributes from two gis datasets that represent the same features into one
+- much of the literature focuses on binary feature matching between disparate data sources
+- we introduce an algorithm that performs line feature matching between two data sources and reports the length of overlap between two features
+- we integrate principles from areal weighted interpolation and geometric conflation to provide an algorithm that matches geometries and performs attribute interpolation for complete and partial matches
+- by de-emphasizing the importance of exacty 1-1 matches and relying on principles of spatial interpolation we can integrate datasets that are not representational of identical phenomena but rather similar ones.
+We remove the emphasis on measures of fit essentially
+- our algorithm is incredible flexible
+  - for example, you can match features that have no overlap if they are within distance tolerance and match the orientation critera to find abutting neighbors 
+  - use results to efficiently narrow down
+
+- Our approach considers all relationships "including one-to-null, null-to-one, one-to-one, one-to-many, many-to-one, and many-to many." (Chehreghan & Abbaspour, 2017, 10.1080/15481603.2017.1338390)
+
+- ^ related to above, we need to document the cardinality of our approach one-to-many there is no 1:1 matching of linestrings (Lei, Ting; Lei, Zhen (2019).)
+ - > Beyond the one‐to‐one case, however, matching becomes more complicated and less well‐defined.
+
+### References
+
+[@morgan_travel_2021]
+[@rawlingson_overlaying_2015]
+[@chehreghan_new_2017]
+[@goodchild_simple_1997]
+[@zhang_methods_nodate]
+[@lei_optimal_2019]
+
+::: {#refs}
+:::

### 2024-02-18 r/R/aggregate.R

@@ -15,7 +15,7 @@
 #' y <- st_transform(st_read(yfp, quiet = TRUE), 27700)[2:3, ]
 #'
 #' matches <- rnet_match(x, y, dist_tolerance = 10, angle_tolerance = 5)
-#' rnet_aggregate_intensive(x, y, matches, value)
+ #' rnet_aggregate_intensive(x, y, matches, value)
 #' rnet_aggregate_extensive(x, y, matches, value)
 #' @rdname aggregate
 rnet_aggregate_extensive <- function(
@@ -80,3 +80,61 @@ rnet_aggregate_intensive <- function(
     ))
 }
 
+
+
+#' @param source data.frame that contains attributes that will be attributed to the target dataset
+#' @param matches results from `rnet_match()`. Contain `i`, `j`, and `shared_len`.
+#' @param extensive_vars character vector of variables names in `source` that will be aggregated to the target as an extensive variable. See Details.
+#' @param intensive_vars character vector of variable names in `source` that will be aggregated to hte target as intensive variables. See Details.
+#' @param source_len default `sf::st_length(source)`. The length of linestrings in the source data set. Used for weighted extensive variables.
+#' @param target_len the lengths of each line in the target dataset. Used for intensive variables.
+#' @export
+rnet_aggregate <- function(
+    source, matches, extensive_vars = NULL, intensive_vars = NULL,
+    source_len = sf::st_length(source),
+    target_len = NULL
+) {
+
+  # target_len cannot be missing when intensive_vars is not null
+  # source_len cannot be missing when extensive_vars is not null
+
+  # create column names with suffixes to be used
+  ext_nms <- paste0(extensive_vars, "_ext")
+  int_nms <- paste0(intensive_vars, "_int")
+
+  # fetch neighbor index
+  j <- matches$j
+
+  # get the neighboring values for each of the intensive and extensive vars
+  # we give them names of the columns with a suffix based on whether or
+  # not it is an extensive or intensive variable
+  ij <- rlang::set_names(
+    lapply(
+      c(extensive_vars, intensive_vars),
+      function(.x) y[[.x]][j]
+    ),
+    c(ext_nms, int_nms)
+  )
+
+  # combine the `ij` values to the original matches
+  dplyr::bind_cols(matches, ij) |>
+    dplyr::mutate(
+      # calculate weight for intensive variables
[…]

### 2024-02-19 r/R/aggregate.R
add categorical variable handling

@@ -90,10 +90,15 @@ rnet_aggregate_intensive <- function(
 #' @param target_len the lengths of each line in the target dataset. Used for intensive variables.
 #' @export
 rnet_aggregate <- function(
-    source, matches, extensive_vars = NULL, intensive_vars = NULL,
+    source,
+    matches,
+    extensive_vars = NULL,
+    intensive_vars = NULL,
+    categorical_vars = NULL,
     source_len = sf::st_length(source),
     target_len = NULL
 ) {
+
 
   # target_len cannot be missing when intensive_vars is not null
   # source_len cannot be missing when extensive_vars is not null
@@ -110,20 +115,32 @@ rnet_aggregate <- function(
   # not it is an extensive or intensive variable
   ij <- rlang::set_names(
     lapply(
-      c(extensive_vars, intensive_vars),
+      c(extensive_vars, intensive_vars, categorical_vars),
       function(.x) y[[.x]][j]
     ),
-    c(ext_nms, int_nms)
+    c(ext_nms, int_nms, categorical_vars)
   )
 
   # combine the `ij` values to the original matches
-  dplyr::bind_cols(matches, ij) |>
+  wts <- dplyr::bind_cols(matches, ij) |>
     dplyr::mutate(
       # calculate weight for intensive variables
       wt_int = shared_len / as.numeric(x_len[i]),
       # calulate weight for extensive variables
       wt_ext = shared_len / as.numeric(y_len[j])
+    )
+
+  # calculate the proportions of categorical variables
+  # for matched lines
+  # FIXME do we apply weighting like Tobler?
+  # https://pysal.org/tobler/generated/tobler.area_weighted.area_interpolate.html#tobler.area_weighted.area_interpolate
+  cat_res <- wts |>
+    dplyr::reframe(
+      dplyr::across(dplyr::all_of(categorical_vars), ~ calc_props(i, .x))
     ) |>
+    tidyr::unnest(dplyr::all_of(categorical_vars))
+
+  numeric_res <- wts |>
     dplyr::group_by(i) |>
     dplyr::summarise(
       # handle extensive variables
@@ -137,4 +154,39 @@ rnet_aggregate <- function(
         ~ weighted.mean(.x, wt_int, na.rm = TRUE)
       )
     )
+
+  bind_cols(numeric_res, cat_res)
+}
+
+
+calc_props <- function(
+    i,
+    var,
+    name_repair = "unique",
[…]

### 2024-02-19 r/R/aggregate.R
fix typos

@@ -81,7 +81,7 @@ rnet_aggregate_intensive <- function(
 }
 
 
-
+#' Aggregate Network Matches
 #' @param source data.frame that contains attributes that will be attributed to the target dataset
 #' @param matches results from `rnet_match()`. Contain `i`, `j`, and `shared_len`.
 #' @param extensive_vars character vector of variables names in `source` that will be aggregated to the target as an extensive variable. See Details.
@@ -125,9 +125,9 @@ rnet_aggregate <- function(
   wts <- dplyr::bind_cols(matches, ij) |>
     dplyr::mutate(
       # calculate weight for intensive variables
-      wt_int = shared_len / as.numeric(x_len[i]),
+      wt_int = shared_len / as.numeric(target_len[i]),
       # calulate weight for extensive variables
-      wt_ext = shared_len / as.numeric(y_len[j])
+      wt_ext = shared_len / as.numeric(source_len[j])
     )
 
   # calculate the proportions of categorical variables

### 2024-02-19 paper.qmd
Update paper.qmd
Co-authored-by: Robin Lovelace <Robinlovelace@users.noreply.github.com>

@@
 
-- use the approximate network matching to create a list of _very strong_ candidate matches with a known shared overlap amount. 
-- The result of the approximate network match can be taken and utilized for vertical conflation using many different approaches 
+- use the approximate network matching to create a list of _strong_ candidate matches with a known shared overlap amount. 
+- The result of the approximate network match can be taken and utilized for 'vertical conflation' using many different approaches 
 - in this paper we introduce a method of spatial interpolation of linear features

### 2024-02-19 paper.qmd
add short section on definitions

@@
 > "Due to the complexity and limitations of existing methods, planners and analysts often have to employ a heavily manual conflation process, which is time‐consuming and often prohibitively expensive." Lei, Ting; Lei, Zhen (2019).
+
+## Definitions
+
+The purpose of this section is to provide clear and concise definitions of terminologies used in the description of this algorithm.
+
+We will use the simple feature access standards definitions of geometric primitives for the sake of consistency.
+In our algorithm we make extensive use of **Line** and **LineString** geometric primitives.
+A Line is defined by two **Point**s that are (x, y) coordinate pairs.
+A LineString is composed of 2 or more Points and "each consecutive pair of Points defines a Line segment." (@sfa FIXME cite).
+Each LineString is referred to as a **feature**.
+An array of features is referred to as a **FeatureCollection** (@geojson_spec).
+
+The objective of this algorithm is to identify matches between two LineString FeatureCollections and measure the length of the match.
+A **match** is a correspondence between features of separate FeatureCollections (@lei_lei_19).
+
+The terminology to refer to the two FeatureCollections is inconsistent.
+For example, Zhang (@zhang_2009) refers to the two FeatureCollections as the **reference** and **target** whereas the Java Conflation Suite (JCS) (@jcs) refers to these as the **reference** and **subject**.
+
+We will use the terms _reference_ FeatureCollection and _subject_ FeatureCollection.
+The objective of the matching algorithm is to match features _from_ the subject _to_ the reference.
+Often the reference is a more detailed and known collection of features.
+
+In the process matching features, **candidate**s are found.
+A candidate is a subject feature that may correspond to a reference feature.
+If a candidate passes tests, it is then deemed a match.
+After matches have been found, attributes from the source feature are often transferred to the reference feature.
+This process of attribute transfer is referred to as **integration**.
 

### 2024-02-20 paper.qmd
update paper.qmd

@@
 
-The terminology to refer to the two FeatureCollections is inconsistent.
-For example, Zhang (@zhang_2009) refers to the two FeatureCollections as the **reference** and **target** whereas the Java Conflation Suite (JCS) (@jcs) refers to these as the **reference** and **subject**.
+The terminology to refer to the two FeatureCollections is inconsistent in the literature.
+For example, Zhang (@zhang_2009) refers to the two FeatureCollections as the reference and target whereas the Java Conflation Suite (JCS) (@jcs) refers to these as the reference and subject.
 
-We will use the terms _reference_ FeatureCollection and _subject_ FeatureCollection.
-The objective of the matching algorithm is to match features _from_ the subject _to_ the reference.
-Often the reference is a more detailed and known collection of features.
+We intentionally lean on terminology used in relational database management systems (RDBMS), instead recognizing a shift towards spatial data science which incorporates RDBMS tools.
+We define the two FeatureCollections as **target** and the **source**.
+The objective of the matching algorithm is to match features _from_ the source _to_ the target.
+Often the target is a more detailed and known collection of features.
 
 In the process matching features, **candidate**s are found.
-A candidate is a subject feature that may correspond to a reference feature.
+A candidate is a source feature that _may_ correspond to a target feature.
 If a candidate passes tests, it is then deemed a match.
-After matches have been found, attributes from the source feature are often transferred to the reference feature.
+After matches have been found, attributes from the source feature are often transferred to the target feature.
 This process of attribute transfer is referred to as **integration**.
@@
 
+- a feature in the source and target may only be matched a maximum of once. 
+- the shared length between i and j may exceeed the total length of i or j. This occurs in parallel ways and larger distance tolerances
+
 ### References

### 2024-02-25 paper.qmd
add further notes to help think this process through more and more

@@
 ---
-title: "Approximate Network Matching"
+title: "ANIME"
+subtitle: "Approximate Network Matching and Integration Enrichment"
 author: 
@@
 
+- datasets are derived from remotely sensed imagery (samal et al., 2004)
+  - this is happening on a much larger scale for example google building footprints (https://sites.research.google/open-buildings/)
+  - features may not always be complete or entirely accurate but can be used to supplement missing data or address changes in a dataset
+
+- matching generalized to more detailed geometry is tough and an unsolved problem , see "LoD" mentions (Ick-Hoi Kim, Chen-Chieh Feng & Yi-Chen Wang), (Mustiere & Devogele, 2008), (Zhang 2009)
+
+ - In a road database, several objects may represent different parts of the same real-world road, e.g., each lane in a highway could be represented by a different object,(Zhang 2009) 
+  - some features can be have no corresponding features or many
+
+- missing join keys between disparate datasets or missing semantic information requires a geometric only approach (Ick-Hoi Kim, Chen-Chieh Feng & Yi-Chen Wang, )
+
+> "a simple overlay of the sources would not automatically reveal correspondence." (Samal et al, 2004)
+
 ## Definitions
@@
 
-We intentionally lean on terminology used in relational database management systems (RDBMS), instead recognizing a shift towards spatial data science which incorporates RDBMS tools.
-We define the two FeatureCollections as **target** and the **source**.
+Instead, we use define the two FeatureCollections as the **source** and the **target** as utilized by Comber and Zeng's work on areal interpolation (@comber_zeng_2019).
+The source and target terminology is also much more commonly used in relational database management systems (RDBMS) and in the wider data science ecosystem.
+The use of these terms is an intentional recognition in the shift towards a more general field of spatial data science.
 The objective of the matching algorithm is to match features _from_ the source _to_ the target.
@@
 After matches have been found, attributes from the source feature are often transferred to the target feature.
-This process of attribute transfer is referred to as **integration**.
+This process of attribute transfer is referred to as **integration** (FIXME CITE).
 
 ## Existing algorithms
+
+- there is a vast literature dedicated to matching linestring features between datasets.
+These can be thought of a category of binary classification algorithms that identify if a feature in the source FeatureCollection is a match to the target.
+- the classification is often based on geometric or semantic criteria, or both.
+The true challenge lies in the absence of semantic information.
+We therefore focus on matching strictly based on geometric criteria.
+- most commonly these criteria are location (proximity), orientation (angle), length, and shape similarity (often measured by some type of distance e.g.
+freshet, hausdorff or avg.
+euclidean from linestring vertices)
+
+Some choice algorithms. By far not exhaustive. Demonstrate some common unique strategies that are employed
 
@@
 
+- Java Conflation Suite (JCS, 2003)
+
+- delimited strokes algorithm 2009 zhang
+
+- Kim et al., 2017 creates an algorithm to conflate historic road network with new ones
+  - uses a number of spatial similarity measures with c4.5 decision tree to classify 
+  - "linear directional mean" of each line segment in a LineString
+    - challenging for longer linestrings with many segments
+  - shorter line median Hausdorff distance (SMHD)
+  - absolute value of cosine similarity (aCS)
+    - note we do something very similar
+
 - (Chehreghan & Abbaspour, 2018) use a genetic algorithm. These are computational expensive and do not sacle well
 
-- overline approach
+- Esri conflation toolset: buffer analysis and similarity
+
+- overline approach (morgan and lovelace)
 
@@
 - orientation (angle)
-- length
+
+- where we differ is in the use of length
+- most existing algorithms perform matching between the source and target based on a measure of length of each LineString. Common are average euclidean distance, hausdorff, and frechet distance.
+- each of these evaluates the linestrings in their entirety 
+- the algorithms then use the spatial criteria to perform a binary classification of either a match or not
+- we do
 
 ## Algorithm overview
+
+- "a good matching algorithm should be able to create high quality results at a high speed." (Zhang, 2009)
+
+- many algorithms are focused on 1:1 correspondence.
+ours supports `m:n` correspondence.
+Each feature in the target FeatureCollection can be matched as few as 0 times or to every feature in the source FeatureCollection.
+  - Our approach considers all relationships "including one-to-null, null-to-one, one-to-one, one-to-many, many-to-one, and many-to many." (Chehreghan & Abbaspour, 2017, 10.1080/15481603.2017.1338390)
+  - ^ related to above, we need to document the cardinality of our approach one-to-many there is no 1:1 matching of linestrings (Lei, Ting; Lei, Zhen (2019).)
+   - > Beyond the one‐to‐one case, however, matching becomes more complicated and less well‐defined.
+
+- consider our algorithm as a way of identifying potential matches that can be pruned if desired
+  - we do not provide a binary classification but a way to identify correspondence between a target and source and measure the amount of correspondence.
 
@@
 
-## Interpolation of values
+## Integration of Attributes
 
-<!-- ## Applications  -->
+- we integrate principles from areal weighted interpolation and geometric conflation to provide an algorithm that matches geometries and performs attribute interpolation for complete and partial matches
+- by de-emphasizing the importance of exacty 1-1 matches and relying on principles of spatial interpolation we can integrate datasets that are not representational of identical phenomena but rather similar ones.
+We remove the emphasis on measures of fit essentially
 
-<!-- - serves as the basis of joins and weighted aggregations -->
-<!-- - NPT scottland  -->
-<!-- - other possible use cases  -->
-<!-- - can be used to subset networks based on shared characteristics e.g. road classification and matched between those subsets  -->
+- in the cases where a complete match is identified this ammounts to a 1:1 attribute transfer
+
+- we rely on intensive and extensive attribute transfer.
+- TODO CITE r-spatial book and Tobler pysal library
+
+- the shared length between i and j may exceed the total length of i or j. This occurs when multiple matches are made between components most often in parallel ways
+
+## Adaptability of Approximate Network Matching
+
+- our algorithm is incredible flexible
+  - for example, you can match features that have no overlap if they are within distance tolerance and match the orientation critera to find abutting neighbors 
+  - use results to efficiently narrow down matching candidates for manual verification
+- can be used as a way to very quickly narrow down matches. Shared overlap can be used to provide a binary classification
 
@@
 
-## Notes
-
-- conflation between two datasets that represent the same area is called "vertical conflation" - JCS technical report
-
-- "version matching" (JCS) features are assumed to be the same and transferring of attributes
-
-- This is a problem of geometric conflation, trying to incorporate attributes from two gis datasets that represent the same features into one
-- much of the literature focuses on binary feature matching between disparate data sources
-- we introduce an algorithm that performs line feature matching between two data sources and reports the length of overlap between two features
-- we integrate principles from areal weighted interpolation and geometric conflation to provide an algorithm that matches geometries and performs attribute interpolation for complete and partial matches
-- by de-emphasizing the importance of exacty 1-1 matches and relying on principles of spatial interpolation we can integrate datasets that are not representational of identical phenomena but rather similar ones.
-We remove the emphasis on measures of fit essentially
-- our algorithm is incredible flexible
-  - for example, you can match features that have no overlap if they are within distance tolerance and match the orientation critera to find abutting neighbors 
-  - use results to efficiently narrow down
-
-- Our approach considers all relationships "including one-to-null, null-to-one, one-to-one, one-to-many, many-to-one, and many-to many." (Chehreghan & Abbaspour, 2017, 10.1080/15481603.2017.1338390)
-
-- ^ related to above, we need to document the cardinality of our approach one-to-many there is no 1:1 matching of linestrings (Lei, Ting; Lei, Zhen (2019).)
- - > Beyond the one‐to‐one case, however, matching becomes more complicated and less well‐defined.
-
-- a feature in the source and target may only be matched a maximum of once. 
-- the shared length between i and j may exceeed the total length of i or j. This occurs in parallel ways and larger distance tolerances
-
 ### References

### 2024-02-25 paper.qmd
add integration formulae

@@
 
-## Integration of Attributes
+## Integration of Numeric Attributes
 
@@
 - the shared length between i and j may exceed the total length of i or j. This occurs when multiple matches are made between components most often in parallel ways
+
+### Extensive Numeric Attribute Integration
+
+sum of shared length / length of y *
+
+let the shared length between target i and source j be the variable $SL_{ij}$
+
+$$
+\hat{Y}_i = \sum_{j} \frac{SL_{ij}}{length(j)} \times Y_j
+$$
+
+### Intensive Numeric Attribute Integration
+
+$$
+\hat{Y}_i = \text{mean}\left(\frac{SL_{ij}}{length(i)} \times Y_j\right)
+$$
+
+## Integration of Categorical Attributes
+
+Similar to the the approach taken for numeric attributes, we can also apply these to categorical variables.
+Say we have a categorical variable $Y$ with $k$ unique values.
+For each $j$ we calculate a frequency table for each $Y_{jk}$.
+We create new variable for each unique $Y_k$.
+$Y_{ik}$ becomes a numeric variable of the frequency.
+$Y_{jk}$ is a dummy variable.
+We now apply the same approach for extensive and intensive attribute integration for each new dummy variable.
+
+### Extensive Categorical Attribute Integration
+
+$\hat{Y_{ik}}$ is equal to the shared length divided by the length of j times $Y_{jk}$ summed for all $ij$ pairs. Note that $Y_{jk}$ can take on only the values of 1 or 0.
+
+$$
+\hat{Y_{ik}} = \sum_{j}{\frac{SL_{ij}}{length(j)}} \times Y_{jk}
+$$
+
+### Intensive Categorical Attribute Integration
+
+$$
+\hat{Y_{ik}} = \text{mean}\left(\frac{SL_{ij}}{length(i)} \times Y_{jk}\right)
+$$
 

### 2024-02-25 r/R/aggregate.R
update aggregate function

@@ -116,7 +116,7 @@ rnet_aggregate <- function(
   ij <- rlang::set_names(
     lapply(
       c(extensive_vars, intensive_vars, categorical_vars),
-      function(.x) y[[.x]][j]
+      function(.x) source[[.x]][j]
     ),
     c(ext_nms, int_nms, categorical_vars)
   )
@@ -155,7 +155,7 @@ rnet_aggregate <- function(
       )
     )
 
-  bind_cols(numeric_res, cat_res)
+  dplyr::bind_cols(numeric_res, cat_res)
 }
 
 

### 2024-03-14 r/data-raw/st_join_example.qmd
Add example file for #35

---
format: gfm
---

This is an example showing how `rnetmatch` can be used with `st_join`,
from issue [\#35](https://github.com/nptscot/rnetmatch/issues/35).

```{r}
library(sf)
library(rnetmatch)
crop_box <- st_bbox(c("xmin" = 427200, xmax = 427500, ymin = 433550, ymax = 433700))

rnet_y <- "https://raw.githubusercontent.com/nptscot/networkmerge/main/data/rnet_armley.geojson" |> 
  read_sf() |>
  # st_geometry() |> 
  st_transform(27700) |> 
  st_crop(crop_box)
rnet_x <- "https://raw.githubusercontent.com/nptscot/networkmerge/main/data/rnet_armley_line.geojson" |>
  read_sf() |> 
  st_crop(crop_box)

# create matches
matches <- rnetmatch::rnet_match(rnet_x, rnet_y, 10, 10)

# function with signature to work with st_join
match_keys <- function(.x = NULL, .y = NULL, matches) {
  dplyr::left_join(tibble::tibble(i = 1:max(matches$i)), matches) |> 
    dplyr::group_by(i) |> 
    dplyr::summarise(j = list(c(j))) |> 
    tibble::deframe() |> 
    lapply(\(.x) {
      if (all(is.na(.x))) {
        integer()
      } else {
        .x
      }
    })
}

x_y_joined_rnetmatch_10_10 <- st_join(
  rnet_x,
  rnet_y,
  join = match_keys,
  # pass matches via dots 
  matches = matches
)
```

### 2024-04-18 r/README.qmd
Add example with negative slope
Update readme

@@
 
-You may also want to install it locally:
+Install it locally with:
 
@@
 #| eval: false
-devtools::install()
+devtools::build()
 ```
@@
 ```{r}
-x = sf::read_sf("data-raw/geojson/princes_street_minimal_x_1.geojson")
+list.files("data-raw/geojson")
+# File is in rnetmatch/r/inst/extdata/princes_street_minimal.geojson
+f = 
+x = sf::read_sf("inst/extdata/princes_street_minimal_x_1.geojson")
 x = x |> sf::st_transform("EPSG:27700")
-y = sf::read_sf("data-raw/geojson/princes_street_minimal.geojson")
+y = sf::read_sf("inst/extdata/princes_street_minimal.geojson")
 y = y |> sf::st_transform("EPSG:27700")
@@
 ```{r}
-matched_df = rnet_match(x, y, dist_tolerance = 20, slope_tolerance = 0.1, trees = "xy")
+matched_df = rnet_match(x, y, dist_tolerance = 20, angle_tolerance = 0.1, trees = "xy")
 y_to_match = y |>
@@
 ```{r}
+x_joined$value
 round(x_joined$value)

### 2024-04-18 r/README.qmd
Update README

@@
 list.files("data-raw/geojson")
-# File is in rnetmatch/r/inst/extdata/princes_street_minimal.geojson
-f = 
 x = sf::read_sf("inst/extdata/princes_street_minimal_x_1.geojson")
@@
 ```{r}
-matched_df = rnet_match(x, y, dist_tolerance = 20, angle_tolerance = 0.1, trees = "xy")
+nrow(x)
+nrow(y)
+matched_df = rnet_match(x, y, dist_tolerance = 20, angle_tolerance = 20, trees = "xy")
+nrow(matched_df)
+matched_df
+length_x = sf::st_length(x) |>
+  # Drop units:
+  units::drop_units()
+length_x
+
 y_to_match = y |>
@@
 matched_df_y = left_join(matched_df, y_to_match)
-x$length_x = sf::st_length(x) |>
-  # Drop units:
-  units::drop_units()
 x_to_match = x |>
@@
 total_distance_y = sum(as.numeric(y$value * sf::st_length(y)))
+total_distance_y
 ```

### 2024-04-23 r/README.qmd

@@
 
-Basic test
+# Basic test
+
+```{r}
+x = sf::st_sf(
+  geometry = sf::st_sfc(
+    sf::st_linestring(matrix(c(1, 1, 5, 1), ncol = 2, byrow = TRUE))
+  )
+)
+
+y = sf::st_sf(
+  geometry = c(
+    sf::st_sfc(
+      sf::st_linestring(matrix(c(2, 2, 6, 2), ncol = 2, byrow = TRUE))
+    ),
+    sf::st_sfc(
+      sf::st_linestring(matrix(c(1, 0.5, 5, 0.5), ncol = 2, byrow = TRUE))
+    )
+  )
+)
+x$object = "x"
+y$object = "y"
+xy = rbind(
+  x,
+  y
+)
+plot(xy)
+plot(x, add = TRUE)
+plot(y)
+```
+
+```{r}
+matched_df = rnet_match(x, y, dist_tolerance = 1, angle_tolerance = 0, trees = "xy")
+matched_df
+```
+
+# Another test
 
